### 2018-03-01 logs/18-02-06.tex

\documentclass{article}
%easy margins
\usepackage[top=.2in, bottom=.4in, left=.8in, right=.8in]{geometry}
%http links in text
\usepackage{hyperref}
%pictures
\usepackage{graphicx}
%quotes
\usepackage{csquotes}
%landscape single pages
\usepackage{lscape}
%inline enumeration
\usepackage[inline]{enumitem}
%for double spaces
\usepackage{setspace}
%****************************************************************
% changing font family
\renewcommand{\familydefault}{\sfdefault} %change font

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Adding title and authors
%\title{Cell classification by applying graphlets on HiC data}
%\author{Behnam Rasoolian  \and Liangliang Xu \and Zheng Zhang}
%\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Chaning font to a local font
%\setmainfont[
%SmallCapsFont = Fontin-SmallCaps.otf,
%BoldFont = Fontin-Bold.otf,
%ItalicFont = Fontin-Italic.otf
%]
%{Fontin.otf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%page title
%\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setstretch{2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%micromanage font size, the first argument is
%%size of the font, the second line distance
\fontsize{13pt}{13pt}\selectfont

%No page numbers.
\pagenumbering{gobble}
%no indentaion for new paragraphs.
\setlength{\parindent}{0pt}
I have downloaded some files from 
\url{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE63525}
there are lots of intrachromosomal data in there, corresponding
to (I think!) 11 cell lines. During the correspondense that I had
with University of Missouri (Mr. Oluwadare), he told me that I have
to used the  GM12878 data set it there, which is much larger than
the original \cite{lieberman2009comprehensive} dataset.
There are two inter- and intra-chromosomal data sets there:

\begin{enumerate}
\fontsize{13pt}{3pt}\selectfont
    \item GSE63525\_GM12878\_primary\_interchromosomal\_contact\_matrices.tar.gz\\
          GSE63525\_GM12878\_primary\_intrachromosomal\_contact\_matrices.tar.gz
      \item GSE63525\_GM12878\_combined\_interchromosomal\_contact\_matrices.tar.gz\\
          GSE63525\_GM12878\_combined\_intrachromosomal\_contact\_matrices.tar.gz
\end{enumerate}
I wonder which ones I should use at the moment.

\section*{Things to consider}
\begin{enumerate}
    \item effect of multiple pearson normalizations
    \item revised SCN algorithm
\end{enumerate}
\end{document}