### 2018-03-08 notebook.tex
add correlation as well as hic data

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{Thresholding\_local}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{Local Thresholding}\label{local-thresholding}

In this notebook, I extract orbits for a signle chromosome in all 4 cell
lines, using a different method of thresholding. In order to do this, I
slide a kernel of size k through each pixel and check if it has a
particular property. If the property is satisified, then the pixel is
set to 1, otherwise it is set to 0. The two properties that I experiment
with are \emph{maximum thresholding} and \emph{normal thresholding}. 1.
\textbf{Maximum Thresholding}: in max thresholding, if the pixel is the
maximum of its neighbors with respect to the kernel, then it is set. 2.
\textbf{Normal Thresholding}: in normal thresholding, if the pixel is
larger that the average of the neighbors then it is set.

first I perform the most lates normalization method that I have found to
be most useful: a combination of gaussian normalization, SCN and pearson
correlation, in the same order mentioned. The normalized results are of
type \texttt{pearson(scn(gaussian(graph)))}.

Then I threshold the results by setting pearson values larger than 0 to
1 and the rest to 0. In order to extract graphlets, I used the
\emph{orca} package in R. the script \texttt{./rscript.r} read a binary
thresholded matrix representing an unweighted graph and outputs orbtis 0
through 72 for each node in the network. After thredholding the
matrices, I save them so that orca in R can read it and extract orbits.

By repeating the procedure explained above for all 23 chromosomes, we
will have graphlets for all contact maps.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k+kn}{as} \PY{n+nn}{np}
        \PY{k+kn}{import} \PY{n+nn}{cv2}
        \PY{k+kn}{from} \PY{n+nn}{utility} \PY{k+kn}{import} \PY{o}{*}
        \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
        \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n}{normalization}
        \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n+nb}{filter}
        \PY{k+kn}{import} \PY{n+nn}{os}
        \PY{o}{\PYZpc{}}\PY{k}{load\PYZus{}ext} autoreload
        \PY{o}{\PYZpc{}}\PY{k}{autoreload} 2
        \PY{o}{\PYZpc{}}\PY{k}{pylab} inline
        \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
The autoreload extension is already loaded. To reload it, use:
  \%reload\_ext autoreload
Populating the interactive namespace from numpy and matplotlib

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/watson/scripts/virt/local/lib/python2.7/site-packages/IPython/core/magics/pylab.py:161: UserWarning: pylab import has clobbered these variables: ['all', 'colors', 'log']
`\%matplotlib` prevents importing * from pylab and numpy
  "\textbackslash{}n`\%matplotlib` prevents importing * from pylab and numpy"

    \end{Verbatim}

    \subsection{Initiating matrices}\label{initiating-matrices}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}18}]:} \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{2}
         \PY{n}{chr2} \PY{o}{=} \PY{n}{chr1}
         \PY{k}{print} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{chromosome 1: }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{, chromosome 2: }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
         \PY{c+c1}{\PYZsh{}print observed\PYZus{}mit\PYZus{}original.shape}

         \PY{n}{mit} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../data/extracted\PYZus{}all\PYZus{}contact\PYZus{}MIT/chr\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}500kb.mat}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PYZbs{}
                       \PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
         \PY{n}{call4} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../data/extracted\PYZus{}all\PYZus{}contact\PYZus{}call4/chr\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}500kb.mat}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PYZbs{}
                         \PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
         \PY{n}{rl} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../data/extracted\PYZus{}all\PYZus{}contact\PYZus{}RL/chr\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}500kb.mat}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

         \PY{n+nb}{all} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../data/extracted\PYZus{}all\PYZus{}contact\PYZus{}ALL/chr\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}500kb.mat}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}t}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

         \PY{c+c1}{\PYZsh{}There are some blank rows and columns in the matrix. let\PYZsq{}s remove them.}
         \PY{n}{blankRows0}\PY{p}{,} \PY{n}{blankCols0} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{mit}\PY{p}{)}
         \PY{n}{blankRows1}\PY{p}{,} \PY{n}{blankCols1} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n+nb}{all}\PY{p}{)}
         \PY{n}{blankRows2}\PY{p}{,} \PY{n}{blankCols2} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{rl}\PY{p}{)}
         \PY{n}{blankRows3}\PY{p}{,} \PY{n}{blankCols3} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{call4}\PY{p}{)}
         \PY{n}{blankRows} \PY{o}{=} \PY{n}{Set}\PY{p}{(}\PY{p}{[}\PY{p}{]}\PY{p}{)}
         \PY{n}{blankCols} \PY{o}{=} \PY{n}{Set}\PY{p}{(}\PY{p}{[}\PY{p}{]}\PY{p}{)}
         \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows0}\PY{p}{)}
         \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows1}\PY{p}{)}
         \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows2}\PY{p}{)}
         \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows3}\PY{p}{)}
         \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols0}\PY{p}{)}
         \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols1}\PY{p}{)}
         \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols2}\PY{p}{)}
         \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols3}\PY{p}{)}
         \PY{n}{mit} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{mit}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
         \PY{n+nb}{all} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n+nb}{all}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
         \PY{n}{rl} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{rl}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
         \PY{n}{call4} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{call4}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
chromosome 1: 2, chromosome 2: 2
file directory: ../data/extracted\_all\_contact\_MIT/chr\_02\_02\_500kb.mat
file directory: ../data/extracted\_all\_contact\_call4/chr\_02\_02\_500kb.mat
file directory: ../data/extracted\_all\_contact\_RL/chr\_02\_02\_500kb.mat
file directory: ../data/extracted\_all\_contact\_ALL/chr\_02\_02\_500kb.mat
size of old matrix: (486, 486)
size of new matrix: (480, 480)
size of old matrix: (486, 486)
size of new matrix: (480, 480)
size of old matrix: (486, 486)
size of new matrix: (480, 480)
size of old matrix: (486, 486)
size of new matrix: (480, 480)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}19}]:} \PY{n}{n0}\PY{p}{,} \PY{n}{m0} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{shape}\PY{p}{(}\PY{n}{mit}\PY{p}{)}
         \PY{n}{n1}\PY{p}{,} \PY{n}{m1} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{shape}\PY{p}{(}\PY{n}{call4}\PY{p}{)}
         \PY{n}{n2}\PY{p}{,} \PY{n}{m2} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{shape}\PY{p}{(}\PY{n}{rl}\PY{p}{)}
         \PY{n}{n3}\PY{p}{,} \PY{n}{m3} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{shape}\PY{p}{(}\PY{n+nb}{all}\PY{p}{)}

         \PY{n}{n} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{[}\PY{n}{n0}\PY{p}{,} \PY{n}{m0}\PY{p}{,} \PY{n}{n1}\PY{p}{,} \PY{n}{m1}\PY{p}{,} \PY{n}{n2}\PY{p}{,} \PY{n}{m2}\PY{p}{,} \PY{n}{n3}\PY{p}{,} \PY{n}{m3}\PY{p}{]}\PY{p}{)}

         \PY{n}{images} \PY{o}{=} \PY{p}{[}\PY{n}{mit}\PY{p}{,} \PY{n+nb}{all}\PY{p}{,} \PY{n}{rl}\PY{p}{,} \PY{n}{call4}\PY{p}{]}
         \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of columnts: 2

    \end{Verbatim}

    \begin{verbatim}
<matplotlib.figure.Figure at 0x7f65c3721610>
    \end{verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_4_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}20}]:} \PY{n}{mit} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{n}{scn}\PY{p}{(}\PY{n}{mit}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
         \PY{n+nb}{all} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{n}{scn}\PY{p}{(}\PY{n+nb}{all}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{rl} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{n}{scn}\PY{p}{(}\PY{n}{rl}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
         \PY{n}{call4} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{n}{scn}\PY{p}{(}\PY{n}{call4}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{1}\PY{p}{)} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}21}]:} \PY{n}{images} \PY{o}{=} \PY{p}{[} 
                 \PY{n}{mit}\PY{p}{,} \PY{n+nb}{all}\PY{p}{,} \PY{n}{rl}\PY{p}{,} \PY{n}{call4}
                 \PY{p}{,} \PY{p}{(}\PY{n}{mit} \PY{o}{==} \PY{n+nb}{all}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{mit} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n+nb}{all} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{p}{,} \PY{n+nb}{all} \PY{o}{!=} \PY{n}{mit}
                 \PY{p}{,} \PY{p}{(}\PY{n}{mit} \PY{o}{==} \PY{n}{rl}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{mit} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{rl} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{p}{,} \PY{n}{rl} \PY{o}{!=} \PY{n}{mit}
                 \PY{p}{,} \PY{p}{(}\PY{n}{mit} \PY{o}{==} \PY{n}{call4}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{mit} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{p}{(}\PY{n}{call4} \PY{o}{!=} \PY{l+m+mi}{0}\PY{p}{)} \PY{p}{,} \PY{n}{call4} \PY{o}{!=} \PY{n}{mit}
                  \PY{p}{]} 
         \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{5}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of columnts: 2

    \end{Verbatim}

    \begin{verbatim}
<matplotlib.figure.Figure at 0x7f65c3718390>
    \end{verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_6_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}22}]:} \PY{n+nb}{min}\PY{p}{,} \PY{n+nb}{max}\PY{p}{,} \PY{n}{mean\PYZus{}mit}\PY{p}{,} \PY{n}{std\PYZus{}mit} \PY{o}{=} \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{mit}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
         \PY{n+nb}{min}\PY{p}{,} \PY{n+nb}{max}\PY{p}{,} \PY{n}{mean\PYZus{}rl}\PY{p}{,} \PY{n}{std\PYZus{}rl} \PY{o}{=} \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{rl}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{rl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
         \PY{n+nb}{min}\PY{p}{,} \PY{n+nb}{max}\PY{p}{,} \PY{n}{mean\PYZus{}call4}\PY{p}{,} \PY{n}{std\PYZus{}call4} \PY{o}{=} \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{call4}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{call4}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
         \PY{n+nb}{min}\PY{p}{,} \PY{n+nb}{max}\PY{p}{,} \PY{n}{mean\PYZus{}all}\PY{p}{,} \PY{n}{std\PYZus{}all} \PY{o}{=} \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n+nb}{all}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{all}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
\#-----------------------------------------------------------------------\#
Stats for mit: 
max: 1.000000 | min: 0.000000 | median: 0.000000 | mean: 0.188247 | std: 0.390909
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for rl: 
max: 1.000000 | min: 0.000000 | median: 0.000000 | mean: 0.153229 | std: 0.360208
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for call4: 
max: 1.000000 | min: 0.000000 | median: 0.000000 | mean: 0.158854 | std: 0.365540
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for all: 
max: 1.000000 | min: 0.000000 | median: 0.000000 | mean: 0.159688 | std: 0.366316
\#-----------------------------------------------------------------------\#

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}23}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{images} \PY{o}{=} \PY{p}{[}
                 \PY{p}{(}\PY{n}{mit}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1e\PYZhy{}2}
                 \PY{p}{,} \PY{p}{(}\PY{n+nb}{all}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1e\PYZhy{}2}
                 \PY{p}{,} \PY{p}{(}\PY{n}{rl}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1e\PYZhy{}2}
                 \PY{p}{,} \PY{p}{(}\PY{n}{call4}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1e\PYZhy{}2}
                 \PY{p}{]}
         \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of columnts: 4

    \end{Verbatim}

    \begin{verbatim}
<matplotlib.figure.Figure at 0x7f65c3752790>
    \end{verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_8_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}24}]:} \PY{n}{images} \PY{o}{=} \PY{p}{[}
                 \PY{n}{log}\PY{p}{(}\PY{n}{call4}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}
                 \PY{p}{,} \PY{n}{log}\PY{p}{(}\PY{n}{mit}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}
                 \PY{p}{,} \PY{n}{log}\PY{p}{(}\PY{n}{rl}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}
                 \PY{p}{,} \PY{n}{log}\PY{p}{(}\PY{n+nb}{all}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}
                 \PY{p}{]}
         \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of columnts: 4

    \end{Verbatim}

    \begin{verbatim}
<matplotlib.figure.Figure at 0x7f65c37400d0>
    \end{verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_9_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}10}]:} \PY{n}{convertBinaryMatToOrcaReadable}\PY{p}{(}\PY{n}{call4} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}call4.edges}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{)}
         \PY{n}{convertBinaryMatToOrcaReadable}\PY{p}{(}\PY{n}{mit} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}mit.edges}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{)}
         \PY{n}{convertBinaryMatToOrcaReadable}\PY{p}{(}\PY{n}{rl} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}rl.edges}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{)}
         \PY{n}{convertBinaryMatToOrcaReadable}\PY{p}{(}\PY{n+nb}{all} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}all.edges}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    % Add a bibliography block to the postdoc

    \end{document}