### 2018-03-08 notebook.tex
added papers

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{02\_graphlet\_correlation\_Analysis}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{Graphlet Correlation
Analysis}\label{graphlet-correlation-analysis}

Remember, in notebook \emph{Thresholding}, I thresholded contact maps
and saved so that orca can read them.

In this notebook, I perform two actions: 1. I apply \texttt{count5}
procedure in orca to extract 73 orbits for each loci in each chromosome.
After this stage, graphlets will be save with \texttt{.graphlets}
extension in the \texttt{./data} folder.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  I then read the graphlets and find both loci-wise distance and
  graphlet-wise correlations between the four cell lines.
\end{enumerate}

By repeating steps 1 and 2 for all 23 chromosomes, I will have pairwise
MIC values for all orbits of all cell lines. I would eventually be able
to have a data array of shape \texttt{A(23\ *\ 73)}, for each pair of
cell lines where \texttt{A{[}i,\ j{]}} denotes the correlation between
the jth orbit in the ith chromosome.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}102}]:} \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k+kn}{as} \PY{n+nn}{np}
          \PY{k+kn}{import} \PY{n+nn}{cv2}
          \PY{k+kn}{from} \PY{n+nn}{utility} \PY{k+kn}{import} \PY{o}{*}
          \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
          \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n}{normalization}
          \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n+nb}{filter}
          \PY{k+kn}{import} \PY{n+nn}{os}
          \PY{o}{\PYZpc{}}\PY{k}{load\PYZus{}ext} autoreload
          \PY{o}{\PYZpc{}}\PY{k}{autoreload} 2
          \PY{o}{\PYZpc{}}\PY{k}{pylab} inline
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
The autoreload extension is already loaded. To reload it, use:
  \%reload\_ext autoreload
Populating the interactive namespace from numpy and matplotlib

    \end{Verbatim}

    \section{1. Orbit Extraction:}\label{orbit-extraction}

The following R-script reads edge files (\texttt{.edges}) corresponding
to a chromosome under \texttt{data/} directory. It then writes the
resulting orbits into \texttt{data/*.graphlets}.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}2}]:} \PY{c+c1}{\PYZsh{}for chr1 in range(1, 24):}
        \PY{c+c1}{\PYZsh{}    os.system(\PYZsq{}Rscript rscript.r \PYZpc{}s\PYZsq{}\PYZpc{}chr1)}
\end{Verbatim}

    \section{2. Loci-wise Distance:}\label{loci-wise-distance}

I can now read the orbitfiles and put them into a \(N \times O\) matrix
where \(N\) denotes number of loci in the chromosome and \(O\) denotes
the number or obitals extracted, which can be either 14 or 73 (depending
on whether \texttt{count4()} or \texttt{count5()} are called in the
R-script above.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}152}]:} \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{2}
          \PY{n}{graphlets\PYZus{}mit} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}mit.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
          \PY{n}{graphlets\PYZus{}all} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}all.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
          \PY{n}{graphlets\PYZus{}rl} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}rl.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
          \PY{n}{graphlets\PYZus{}call4} \PY{o}{=} \PY{n}{readMat}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}call4.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{n}{graphlets\PYZus{}all}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{n}{graphlets\PYZus{}rl}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{n}{graphlets\PYZus{}call4}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr02\_chr02\_mit.graphlets
file directory: data/chr02\_chr02\_all.graphlets
file directory: data/chr02\_chr02\_rl.graphlets
file directory: data/chr02\_chr02\_call4.graphlets
(478, 73)
(478, 73)
(478, 73)
(478, 73)

    \end{Verbatim}

    Calculating Loci-wise correlations:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}153}]:} \PY{n}{mit\PYZus{}all\PYZus{}corr} \PY{o}{=} \PYZbs{}
              \PY{n}{graphlet\PYZus{}correlational\PYZus{}distance}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{p}{,} \PY{n}{graphlets\PYZus{}all}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}154}]:} \PY{n}{mit\PYZus{}mit\PYZus{}corr} \PY{o}{=} \PYZbs{}
              \PY{n}{graphlet\PYZus{}correlational\PYZus{}distance}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{p}{,} \PY{n}{graphlets\PYZus{}mit}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}155}]:} \PY{n}{mit\PYZus{}call4\PYZus{}corr} \PY{o}{=} \PYZbs{}
              \PY{n}{graphlet\PYZus{}correlational\PYZus{}distance}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{p}{,} \PY{n}{graphlets\PYZus{}call4}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}156}]:} \PY{n}{mit\PYZus{}rl\PYZus{}corr} \PY{o}{=} \PYZbs{}
              \PY{n}{graphlet\PYZus{}correlational\PYZus{}distance}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{p}{,} \PY{n}{graphlets\PYZus{}rl}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}157}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{images} \PY{o}{=} \PY{p}{[} \PY{n}{mit\PYZus{}mit\PYZus{}corr}
                    \PY{p}{,} \PY{n}{mit\PYZus{}all\PYZus{}corr}
                   \PY{p}{,} \PY{n}{mit\PYZus{}rl\PYZus{}corr}
                   \PY{p}{,} \PY{n}{mit\PYZus{}call4\PYZus{}corr}\PY{p}{]}
          \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of columnts: 2

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_11_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}161}]:} \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Total correlation statistics}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{mit\PYZus{}mit\PYZus{}corr}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs mit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{mit\PYZus{}all\PYZus{}corr}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs all}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{mit\PYZus{}rl\PYZus{}corr}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs rl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{mit\PYZus{}call4\PYZus{}corr}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs call4}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+se}{\PYZbs{}n}\PY{l+s+s2}{Total correlation statistics}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}mit\PYZus{}corr}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs mit}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}all\PYZus{}corr}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs all}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}rl\PYZus{}corr}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs rl}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
          \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}call4\PYZus{}corr}\PY{p}{)}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs call4}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Total correlation statistics
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 4.469689 | min: 0.000000 | median: 0.118804 | mean: 0.177694 | std: 0.406987
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 4.469821 | min: 0.000000 | median: 0.254837 | mean: 0.331607 | std: 0.424543
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 4.469689 | min: 0.000000 | median: 0.368116 | mean: 0.429916 | std: 0.418677
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 4.469790 | min: 0.000000 | median: 0.296439 | mean: 0.359209 | std: 0.420711
\#-----------------------------------------------------------------------\#

Total correlation statistics
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 0.000000 | min: 0.000000 | median: 0.000000 | mean: 0.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 1.805932 | min: 0.000000 | median: 0.243699 | mean: 0.287908 | std: 0.178464
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 1.235199 | min: 0.000000 | median: 0.368138 | mean: 0.391644 | std: 0.196214
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 1.392509 | min: 0.000000 | median: 0.298928 | mean: 0.317962 | std: 0.175240
\#-----------------------------------------------------------------------\#

    \end{Verbatim}

\begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}161}]:} (1.3925089365419872, 0.0, 0.3179615758460628, 0.17523973020518074)
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}182}]:} \PY{n}{n} \PY{o}{=} \PY{n}{mit\PYZus{}mit\PYZus{}corr}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
          \PY{c+c1}{\PYZsh{}t1 = (t\PYZus{}test(np.diagonal(mit\PYZus{}mit\PYZus{}corr)))}
          \PY{n}{t2} \PY{o}{=} \PY{p}{(}\PY{n}{t\PYZus{}test}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}all\PYZus{}corr}\PY{p}{)}\PY{p}{)}\PY{p}{)}
          \PY{n}{t3} \PY{o}{=} \PY{p}{(}\PY{n}{t\PYZus{}test}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}rl\PYZus{}corr}\PY{p}{)}\PY{p}{)}\PY{p}{)}
          \PY{n}{t4} \PY{o}{=} \PY{p}{(}\PY{n}{t\PYZus{}test}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}call4\PYZus{}corr}\PY{p}{)}\PY{p}{)}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{n}{n}\PY{p}{)}
          \PY{c+c1}{\PYZsh{}print(t1)}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{t statistic for mit vs all: }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}} \PY{n}{t2}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{t statistic for mit vs rl: }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}} \PY{n}{t3}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{t statistic for mit vs call4: }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}} \PY{n}{t4}\PY{p}{)}
          \PY{k+kn}{from} \PY{n+nn}{scipy.stats} \PY{k+kn}{import} \PY{n}{t}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{p\PYZhy{}value for mit\PYZhy{}all: }\PY{l+s+si}{\PYZpc{}10.3e}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n}{t}\PY{o}{.}\PY{n}{cdf}\PY{p}{(}\PY{n}{t2}\PY{p}{,} \PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{p\PYZhy{}value for mit\PYZhy{}rl: }\PY{l+s+si}{\PYZpc{}10.3e}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n}{t}\PY{o}{.}\PY{n}{cdf}\PY{p}{(}\PY{n}{t3}\PY{p}{,} \PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{)}
          \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{p\PYZhy{}value for mit\PYZhy{}call4: }\PY{l+s+si}{\PYZpc{}10.3e}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n}{t}\PY{o}{.}\PY{n}{cdf}\PY{p}{(}\PY{n}{t4}\PY{p}{,} \PY{n}{n}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
478
t statistic for mit vs all: 18
t statistic for mit vs rl: 19
t statistic for mit vs call4: 19
p-value for mit-all:  0.000e+00
p-value for mit-rl:  0.000e+00
p-value for mit-call4:  0.000e+00

    \end{Verbatim}

    \section{3. Orbit-wise Correlation}\label{orbit-wise-correlation}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}192}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
              \PY{n}{graphlets\PYZus{}mit} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}mit.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}all} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}all.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}rl} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}rl.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}call4} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}call4.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{mit\PYZus{}all\PYZus{}corr\PYZus{}orbit\PYZus{}wise} \PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}all}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}mit\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}call4\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}call4}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}rl\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}rl}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{images} \PY{o}{=} \PY{p}{[} \PY{n}{mit\PYZus{}mit\PYZus{}corr\PYZus{}orbit\PYZus{}wise}
                    \PY{p}{,} \PY{n}{mit\PYZus{}all\PYZus{}corr\PYZus{}orbit\PYZus{}wise}
                   \PY{p}{,} \PY{n}{mit\PYZus{}rl\PYZus{}corr\PYZus{}orbit\PYZus{}wise}
                   \PY{p}{,} \PY{n}{mit\PYZus{}call4\PYZus{}corr\PYZus{}orbit\PYZus{}wise}
                   \PY{p}{]}
              \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr01\_chr01\_mit.graphlets
file directory: data/chr01\_chr01\_all.graphlets
file directory: data/chr01\_chr01\_rl.graphlets
file directory: data/chr01\_chr01\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr02\_chr02\_mit.graphlets
file directory: data/chr02\_chr02\_all.graphlets
file directory: data/chr02\_chr02\_rl.graphlets
file directory: data/chr02\_chr02\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_3.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr03\_chr03\_mit.graphlets
file directory: data/chr03\_chr03\_all.graphlets
file directory: data/chr03\_chr03\_rl.graphlets
file directory: data/chr03\_chr03\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_5.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr04\_chr04\_mit.graphlets
file directory: data/chr04\_chr04\_all.graphlets
file directory: data/chr04\_chr04\_rl.graphlets
file directory: data/chr04\_chr04\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_7.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr05\_chr05\_mit.graphlets
file directory: data/chr05\_chr05\_all.graphlets
file directory: data/chr05\_chr05\_rl.graphlets
file directory: data/chr05\_chr05\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_9.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr06\_chr06\_mit.graphlets
file directory: data/chr06\_chr06\_all.graphlets
file directory: data/chr06\_chr06\_rl.graphlets
file directory: data/chr06\_chr06\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_11.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr07\_chr07\_mit.graphlets
file directory: data/chr07\_chr07\_all.graphlets
file directory: data/chr07\_chr07\_rl.graphlets
file directory: data/chr07\_chr07\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_13.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr08\_chr08\_mit.graphlets
file directory: data/chr08\_chr08\_all.graphlets
file directory: data/chr08\_chr08\_rl.graphlets
file directory: data/chr08\_chr08\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_15.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr09\_chr09\_mit.graphlets
file directory: data/chr09\_chr09\_all.graphlets
file directory: data/chr09\_chr09\_rl.graphlets
file directory: data/chr09\_chr09\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_17.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr10\_chr10\_mit.graphlets
file directory: data/chr10\_chr10\_all.graphlets
file directory: data/chr10\_chr10\_rl.graphlets
file directory: data/chr10\_chr10\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_19.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr11\_chr11\_mit.graphlets
file directory: data/chr11\_chr11\_all.graphlets
file directory: data/chr11\_chr11\_rl.graphlets
file directory: data/chr11\_chr11\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_21.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr12\_chr12\_mit.graphlets
file directory: data/chr12\_chr12\_all.graphlets
file directory: data/chr12\_chr12\_rl.graphlets
file directory: data/chr12\_chr12\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_23.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr13\_chr13\_mit.graphlets
file directory: data/chr13\_chr13\_all.graphlets
file directory: data/chr13\_chr13\_rl.graphlets
file directory: data/chr13\_chr13\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_25.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr14\_chr14\_mit.graphlets
file directory: data/chr14\_chr14\_all.graphlets
file directory: data/chr14\_chr14\_rl.graphlets
file directory: data/chr14\_chr14\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_27.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr15\_chr15\_mit.graphlets
file directory: data/chr15\_chr15\_all.graphlets
file directory: data/chr15\_chr15\_rl.graphlets
file directory: data/chr15\_chr15\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_29.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr16\_chr16\_mit.graphlets
file directory: data/chr16\_chr16\_all.graphlets
file directory: data/chr16\_chr16\_rl.graphlets
file directory: data/chr16\_chr16\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_31.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr17\_chr17\_mit.graphlets
file directory: data/chr17\_chr17\_all.graphlets
file directory: data/chr17\_chr17\_rl.graphlets
file directory: data/chr17\_chr17\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_33.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr18\_chr18\_mit.graphlets
file directory: data/chr18\_chr18\_all.graphlets
file directory: data/chr18\_chr18\_rl.graphlets
file directory: data/chr18\_chr18\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_35.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr19\_chr19\_mit.graphlets
file directory: data/chr19\_chr19\_all.graphlets
file directory: data/chr19\_chr19\_rl.graphlets
file directory: data/chr19\_chr19\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_37.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr20\_chr20\_mit.graphlets
file directory: data/chr20\_chr20\_all.graphlets
file directory: data/chr20\_chr20\_rl.graphlets
file directory: data/chr20\_chr20\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_39.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr21\_chr21\_mit.graphlets
file directory: data/chr21\_chr21\_all.graphlets
file directory: data/chr21\_chr21\_rl.graphlets
file directory: data/chr21\_chr21\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_41.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr22\_chr22\_mit.graphlets
file directory: data/chr22\_chr22\_all.graphlets
file directory: data/chr22\_chr22\_rl.graphlets
file directory: data/chr22\_chr22\_call4.graphlets
Number of columns (68 and 67) not the same.
            picking the minimum value{\ldots}
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_43.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr23\_chr23\_mit.graphlets
file directory: data/chr23\_chr23\_all.graphlets
file directory: data/chr23\_chr23\_rl.graphlets
file directory: data/chr23\_chr23\_call4.graphlets
Number of columnts: 4

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_45.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}198}]:} \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
              \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}mit} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}mit.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}all} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}all.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}rl} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}rl.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{graphlets\PYZus{}call4} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}call4.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{mit\PYZus{}all\PYZus{}corr\PYZus{}orbit\PYZus{}wise} \PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}all}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}mit\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}call4\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}call4}\PY{o}{.}\PY{n}{T}\PY{p}{)}
              \PY{n}{mit\PYZus{}rl\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{o}{=} \PYZbs{}
                  \PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets\PYZus{}mit}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets\PYZus{}rl}\PY{o}{.}\PY{n}{T}\PY{p}{)} 
              \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}mit\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit vs mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}all\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit vs all}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}rl\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit vs rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
              \PY{n}{print\PYZus{}statistics}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{mit\PYZus{}call4\PYZus{}corr\PYZus{}orbit\PYZus{}wise}\PY{p}{)}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit vs call4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Chromosome 1
file directory: data/chr01\_chr01\_mit.graphlets
file directory: data/chr01\_chr01\_all.graphlets
file directory: data/chr01\_chr01\_rl.graphlets
file directory: data/chr01\_chr01\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.691546 | min: 0.264693 | median: 0.447225 | mean: 0.458319 | std: 0.107821
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.488262 | min: 0.005228 | median: 0.064152 | mean: 0.115254 | std: 0.145756
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.735380 | min: 0.067395 | median: 0.209633 | mean: 0.266105 | std: 0.183532
\#-----------------------------------------------------------------------\#
Chromosome 2
file directory: data/chr02\_chr02\_mit.graphlets
file directory: data/chr02\_chr02\_all.graphlets
file directory: data/chr02\_chr02\_rl.graphlets
file directory: data/chr02\_chr02\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.612056 | min: 0.154044 | median: 0.397228 | mean: 0.394162 | std: 0.112857
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.544215 | min: 0.111064 | median: 0.186193 | mean: 0.239152 | std: 0.127777
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.654175 | min: 0.179010 | median: 0.387024 | mean: 0.398918 | std: 0.120441
\#-----------------------------------------------------------------------\#
Chromosome 3
file directory: data/chr03\_chr03\_mit.graphlets
file directory: data/chr03\_chr03\_all.graphlets
file directory: data/chr03\_chr03\_rl.graphlets
file directory: data/chr03\_chr03\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.662087 | min: 0.104050 | median: 0.202309 | mean: 0.256182 | std: 0.150717
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.547617 | min: -0.099389 | median: -0.048790 | mean: 0.040852 | std: 0.186404
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.612281 | min: 0.211936 | median: 0.284129 | mean: 0.317399 | std: 0.101562
\#-----------------------------------------------------------------------\#
Chromosome 4
file directory: data/chr04\_chr04\_mit.graphlets
file directory: data/chr04\_chr04\_all.graphlets
file directory: data/chr04\_chr04\_rl.graphlets
file directory: data/chr04\_chr04\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.671499 | min: 0.122204 | median: 0.288221 | mean: 0.322727 | std: 0.126746
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.594044 | min: 0.010192 | median: 0.105796 | mean: 0.158603 | std: 0.167435
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.619095 | min: 0.180210 | median: 0.219814 | mean: 0.274621 | std: 0.124545
\#-----------------------------------------------------------------------\#
Chromosome 5
file directory: data/chr05\_chr05\_mit.graphlets
file directory: data/chr05\_chr05\_all.graphlets
file directory: data/chr05\_chr05\_rl.graphlets
file directory: data/chr05\_chr05\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.690312 | min: 0.150168 | median: 0.270318 | mean: 0.325072 | std: 0.139688
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.656837 | min: 0.092786 | median: 0.184570 | mean: 0.243654 | std: 0.165851
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.695475 | min: 0.050792 | median: 0.212408 | mean: 0.271915 | std: 0.163642
\#-----------------------------------------------------------------------\#
Chromosome 6
file directory: data/chr06\_chr06\_mit.graphlets
file directory: data/chr06\_chr06\_all.graphlets
file directory: data/chr06\_chr06\_rl.graphlets
file directory: data/chr06\_chr06\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.632833 | min: 0.057611 | median: 0.200461 | mean: 0.261235 | std: 0.151247
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.603632 | min: 0.085527 | median: 0.163064 | mean: 0.221928 | std: 0.146656
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.641864 | min: 0.146304 | median: 0.345319 | mean: 0.373079 | std: 0.107218
\#-----------------------------------------------------------------------\#
Chromosome 7
file directory: data/chr07\_chr07\_mit.graphlets
file directory: data/chr07\_chr07\_all.graphlets
file directory: data/chr07\_chr07\_rl.graphlets
file directory: data/chr07\_chr07\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.610315 | min: 0.112599 | median: 0.257948 | mean: 0.299058 | std: 0.110721
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.557602 | min: 0.093335 | median: 0.173999 | mean: 0.218020 | std: 0.131495
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.624037 | min: 0.194459 | median: 0.260055 | mean: 0.310916 | std: 0.116809
\#-----------------------------------------------------------------------\#
Chromosome 8
file directory: data/chr08\_chr08\_mit.graphlets
file directory: data/chr08\_chr08\_all.graphlets
file directory: data/chr08\_chr08\_rl.graphlets
file directory: data/chr08\_chr08\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.622792 | min: 0.064770 | median: 0.218786 | mean: 0.241913 | std: 0.149463
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.548060 | min: -0.072579 | median: 0.022616 | mean: 0.081929 | std: 0.183618
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.565263 | min: 0.029225 | median: 0.136874 | mean: 0.176655 | std: 0.141294
\#-----------------------------------------------------------------------\#
Chromosome 9
file directory: data/chr09\_chr09\_mit.graphlets
file directory: data/chr09\_chr09\_all.graphlets
file directory: data/chr09\_chr09\_rl.graphlets
file directory: data/chr09\_chr09\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.673587 | min: 0.215801 | median: 0.486040 | mean: 0.475922 | std: 0.089530
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.584072 | min: 0.085406 | median: 0.163250 | mean: 0.193730 | std: 0.114805
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.564110 | min: 0.239796 | median: 0.353665 | mean: 0.358896 | std: 0.070965
\#-----------------------------------------------------------------------\#
Chromosome 10
file directory: data/chr10\_chr10\_mit.graphlets
file directory: data/chr10\_chr10\_all.graphlets
file directory: data/chr10\_chr10\_rl.graphlets
file directory: data/chr10\_chr10\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.580837 | min: 0.039492 | median: 0.149594 | mean: 0.209214 | std: 0.152388
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.610226 | min: 0.028290 | median: 0.136378 | mean: 0.188366 | std: 0.170261
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.557048 | min: 0.128177 | median: 0.192862 | mean: 0.248928 | std: 0.121032
\#-----------------------------------------------------------------------\#
Chromosome 11
file directory: data/chr11\_chr11\_mit.graphlets
file directory: data/chr11\_chr11\_all.graphlets
file directory: data/chr11\_chr11\_rl.graphlets
file directory: data/chr11\_chr11\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.653530 | min: 0.071174 | median: 0.234628 | mean: 0.264804 | std: 0.146381
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.617074 | min: 0.053525 | median: 0.245710 | mean: 0.267019 | std: 0.119514
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.655354 | min: 0.185627 | median: 0.264436 | mean: 0.307051 | std: 0.116659
\#-----------------------------------------------------------------------\#
Chromosome 12
file directory: data/chr12\_chr12\_mit.graphlets
file directory: data/chr12\_chr12\_all.graphlets
file directory: data/chr12\_chr12\_rl.graphlets
file directory: data/chr12\_chr12\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.614255 | min: 0.095773 | median: 0.199803 | mean: 0.247623 | std: 0.146455
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.582140 | min: 0.086246 | median: 0.208719 | mean: 0.257760 | std: 0.132487
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.697678 | min: 0.002999 | median: 0.176178 | mean: 0.236578 | std: 0.164529
\#-----------------------------------------------------------------------\#
Chromosome 13
file directory: data/chr13\_chr13\_mit.graphlets
file directory: data/chr13\_chr13\_all.graphlets
file directory: data/chr13\_chr13\_rl.graphlets
file directory: data/chr13\_chr13\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.656408 | min: -0.096631 | median: 0.156068 | mean: 0.209380 | std: 0.172578
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.595335 | min: -0.013590 | median: 0.111875 | mean: 0.164435 | std: 0.150826
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.619183 | min: -0.053744 | median: 0.118303 | mean: 0.185206 | std: 0.183478
\#-----------------------------------------------------------------------\#
Chromosome 14
file directory: data/chr14\_chr14\_mit.graphlets
file directory: data/chr14\_chr14\_all.graphlets
file directory: data/chr14\_chr14\_rl.graphlets
file directory: data/chr14\_chr14\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.543247 | min: -0.018916 | median: 0.161567 | mean: 0.193250 | std: 0.150059
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.632834 | min: 0.044777 | median: 0.318658 | mean: 0.345974 | std: 0.107771
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.589953 | min: -0.015215 | median: 0.146438 | mean: 0.199090 | std: 0.139573
\#-----------------------------------------------------------------------\#
Chromosome 15
file directory: data/chr15\_chr15\_mit.graphlets
file directory: data/chr15\_chr15\_all.graphlets
file directory: data/chr15\_chr15\_rl.graphlets
file directory: data/chr15\_chr15\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.627744 | min: -0.082786 | median: 0.270007 | mean: 0.295242 | std: 0.137112
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.563995 | min: -0.032939 | median: 0.066092 | mean: 0.130166 | std: 0.169149
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.555546 | min: -0.105676 | median: 0.049453 | mean: 0.122544 | std: 0.172953
\#-----------------------------------------------------------------------\#
Chromosome 16
file directory: data/chr16\_chr16\_mit.graphlets
file directory: data/chr16\_chr16\_all.graphlets
file directory: data/chr16\_chr16\_rl.graphlets
file directory: data/chr16\_chr16\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.663765 | min: 0.069383 | median: 0.262083 | mean: 0.282577 | std: 0.131222
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.571480 | min: 0.110550 | median: 0.250025 | mean: 0.273534 | std: 0.103604
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.512031 | min: -0.025588 | median: 0.119995 | mean: 0.155271 | std: 0.129841
\#-----------------------------------------------------------------------\#
Chromosome 17
file directory: data/chr17\_chr17\_mit.graphlets
file directory: data/chr17\_chr17\_all.graphlets
file directory: data/chr17\_chr17\_rl.graphlets
file directory: data/chr17\_chr17\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.630219 | min: 0.002755 | median: 0.203456 | mean: 0.247660 | std: 0.158113
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.594223 | min: -0.041055 | median: 0.090041 | mean: 0.145245 | std: 0.165721
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.613410 | min: 0.025560 | median: 0.129442 | mean: 0.185947 | std: 0.150671
\#-----------------------------------------------------------------------\#
Chromosome 18
file directory: data/chr18\_chr18\_mit.graphlets
file directory: data/chr18\_chr18\_all.graphlets
file directory: data/chr18\_chr18\_rl.graphlets
file directory: data/chr18\_chr18\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.523809 | min: -0.090717 | median: 0.084290 | mean: 0.116331 | std: 0.151788
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.610116 | min: -0.160623 | median: 0.020117 | mean: 0.101178 | std: 0.192265
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.474894 | min: -0.019191 | median: 0.119772 | mean: 0.158747 | std: 0.116521
\#-----------------------------------------------------------------------\#
Chromosome 19
file directory: data/chr19\_chr19\_mit.graphlets
file directory: data/chr19\_chr19\_all.graphlets
file directory: data/chr19\_chr19\_rl.graphlets
file directory: data/chr19\_chr19\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.635402 | min: 0.028820 | median: 0.258356 | mean: 0.271281 | std: 0.128827
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.529379 | min: -0.104304 | median: 0.157916 | mean: 0.191153 | std: 0.144419
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
Chromosome 20
file directory: data/chr20\_chr20\_mit.graphlets
file directory: data/chr20\_chr20\_all.graphlets
file directory: data/chr20\_chr20\_rl.graphlets
file directory: data/chr20\_chr20\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.575553 | min: -0.031425 | median: 0.167951 | mean: 0.175358 | std: 0.143634
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.570305 | min: -0.011435 | median: 0.236590 | mean: 0.256076 | std: 0.105848
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.553739 | min: 0.094144 | median: 0.240062 | mean: 0.262793 | std: 0.089579
\#-----------------------------------------------------------------------\#
Chromosome 21
file directory: data/chr21\_chr21\_mit.graphlets
file directory: data/chr21\_chr21\_all.graphlets
file directory: data/chr21\_chr21\_rl.graphlets
file directory: data/chr21\_chr21\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
Chromosome 22
file directory: data/chr22\_chr22\_mit.graphlets
file directory: data/chr22\_chr22\_all.graphlets
file directory: data/chr22\_chr22\_rl.graphlets
file directory: data/chr22\_chr22\_call4.graphlets
Number of columns (68 and 67) not the same.
            picking the minimum value{\ldots}
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: nan | min: nan | median: nan | mean: nan | std: nan
\#-----------------------------------------------------------------------\#
Chromosome 23
file directory: data/chr23\_chr23\_mit.graphlets
file directory: data/chr23\_chr23\_all.graphlets
file directory: data/chr23\_chr23\_rl.graphlets
file directory: data/chr23\_chr23\_call4.graphlets
\#-----------------------------------------------------------------------\#
Stats for mit vs mit: 
max: 1.000000 | min: 1.000000 | median: 1.000000 | mean: 1.000000 | std: 0.000000
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs all: 
max: 0.661035 | min: 0.007166 | median: 0.193827 | mean: 0.258310 | std: 0.167515
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs rl: 
max: 0.520688 | min: 0.009042 | median: 0.106761 | mean: 0.168885 | std: 0.144671
\#-----------------------------------------------------------------------\#
\#-----------------------------------------------------------------------\#
Stats for mit vs call4: 
max: 0.594355 | min: 0.094710 | median: 0.184106 | mean: 0.239780 | std: 0.128988
\#-----------------------------------------------------------------------\#

    \end{Verbatim}

    Now Let's repeat the above procedure for all 23 chromosomes and store
them in a dictionary:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}115}]:} \PY{n}{cells} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{all}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{call4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
          \PY{n}{graphlets} \PY{o}{=} \PY{p}{[} \PY{n+nb+bp}{None} \PY{p}{]} \PY{o}{*} \PY{l+m+mi}{24}
          \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
              \PY{n}{graphlets}\PY{p}{[}\PY{n}{chr1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
              \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{n}{graphlets}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
              \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{data/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{cell}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}

          \PY{n}{correlations} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
          \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
              \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
              \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                  \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{n}{cell2}\PY{p}{:}
                      \PY{k}{continue}
                  \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{l+m+mi}{24}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}\PY{p}{)}
                  \PY{c+c1}{\PYZsh{}cell (i, j) will store correlation between orbital j in }
                  \PY{c+c1}{\PYZsh{}ith chromosome}

          \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
              \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{:}
                  \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                      \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{,} \PY{p}{:}\PY{p}{]} \PY{o}{=} \PYZbs{}
                          \PY{n}{np}\PY{o}{.}\PY{n}{diagonal}\PY{p}{(}\PY{n}{pearson}\PY{p}{(}\PY{n}{graphlets}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{o}{.}\PY{n}{T}\PY{p}{,} \PY{n}{graphlets}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{o}{.}\PY{n}{T}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
file directory: data/chr01\_chr01\_mit.graphlets
file directory: data/chr01\_chr01\_all.graphlets
file directory: data/chr01\_chr01\_rl.graphlets
file directory: data/chr01\_chr01\_call4.graphlets
file directory: data/chr02\_chr02\_mit.graphlets
file directory: data/chr02\_chr02\_all.graphlets
file directory: data/chr02\_chr02\_rl.graphlets
file directory: data/chr02\_chr02\_call4.graphlets
file directory: data/chr03\_chr03\_mit.graphlets
file directory: data/chr03\_chr03\_all.graphlets
file directory: data/chr03\_chr03\_rl.graphlets
file directory: data/chr03\_chr03\_call4.graphlets
file directory: data/chr04\_chr04\_mit.graphlets
file directory: data/chr04\_chr04\_all.graphlets
file directory: data/chr04\_chr04\_rl.graphlets
file directory: data/chr04\_chr04\_call4.graphlets
file directory: data/chr05\_chr05\_mit.graphlets
file directory: data/chr05\_chr05\_all.graphlets
file directory: data/chr05\_chr05\_rl.graphlets
file directory: data/chr05\_chr05\_call4.graphlets
file directory: data/chr06\_chr06\_mit.graphlets
file directory: data/chr06\_chr06\_all.graphlets
file directory: data/chr06\_chr06\_rl.graphlets
file directory: data/chr06\_chr06\_call4.graphlets
file directory: data/chr07\_chr07\_mit.graphlets
file directory: data/chr07\_chr07\_all.graphlets
file directory: data/chr07\_chr07\_rl.graphlets
file directory: data/chr07\_chr07\_call4.graphlets
file directory: data/chr08\_chr08\_mit.graphlets
file directory: data/chr08\_chr08\_all.graphlets
file directory: data/chr08\_chr08\_rl.graphlets
file directory: data/chr08\_chr08\_call4.graphlets
file directory: data/chr09\_chr09\_mit.graphlets
file directory: data/chr09\_chr09\_all.graphlets
file directory: data/chr09\_chr09\_rl.graphlets
file directory: data/chr09\_chr09\_call4.graphlets
file directory: data/chr10\_chr10\_mit.graphlets
file directory: data/chr10\_chr10\_all.graphlets
file directory: data/chr10\_chr10\_rl.graphlets
file directory: data/chr10\_chr10\_call4.graphlets
file directory: data/chr11\_chr11\_mit.graphlets
file directory: data/chr11\_chr11\_all.graphlets
file directory: data/chr11\_chr11\_rl.graphlets
file directory: data/chr11\_chr11\_call4.graphlets
file directory: data/chr12\_chr12\_mit.graphlets
file directory: data/chr12\_chr12\_all.graphlets
file directory: data/chr12\_chr12\_rl.graphlets
file directory: data/chr12\_chr12\_call4.graphlets
file directory: data/chr13\_chr13\_mit.graphlets
file directory: data/chr13\_chr13\_all.graphlets
file directory: data/chr13\_chr13\_rl.graphlets
file directory: data/chr13\_chr13\_call4.graphlets
file directory: data/chr14\_chr14\_mit.graphlets
file directory: data/chr14\_chr14\_all.graphlets
file directory: data/chr14\_chr14\_rl.graphlets
file directory: data/chr14\_chr14\_call4.graphlets
file directory: data/chr15\_chr15\_mit.graphlets
file directory: data/chr15\_chr15\_all.graphlets
file directory: data/chr15\_chr15\_rl.graphlets
file directory: data/chr15\_chr15\_call4.graphlets
file directory: data/chr16\_chr16\_mit.graphlets
file directory: data/chr16\_chr16\_all.graphlets
file directory: data/chr16\_chr16\_rl.graphlets
file directory: data/chr16\_chr16\_call4.graphlets
file directory: data/chr17\_chr17\_mit.graphlets
file directory: data/chr17\_chr17\_all.graphlets
file directory: data/chr17\_chr17\_rl.graphlets
file directory: data/chr17\_chr17\_call4.graphlets
file directory: data/chr18\_chr18\_mit.graphlets
file directory: data/chr18\_chr18\_all.graphlets
file directory: data/chr18\_chr18\_rl.graphlets
file directory: data/chr18\_chr18\_call4.graphlets
file directory: data/chr19\_chr19\_mit.graphlets
file directory: data/chr19\_chr19\_all.graphlets
file directory: data/chr19\_chr19\_rl.graphlets
file directory: data/chr19\_chr19\_call4.graphlets
file directory: data/chr20\_chr20\_mit.graphlets
file directory: data/chr20\_chr20\_all.graphlets
file directory: data/chr20\_chr20\_rl.graphlets
file directory: data/chr20\_chr20\_call4.graphlets
file directory: data/chr21\_chr21\_mit.graphlets
file directory: data/chr21\_chr21\_all.graphlets
file directory: data/chr21\_chr21\_rl.graphlets
file directory: data/chr21\_chr21\_call4.graphlets
file directory: data/chr22\_chr22\_mit.graphlets
file directory: data/chr22\_chr22\_all.graphlets
file directory: data/chr22\_chr22\_rl.graphlets
file directory: data/chr22\_chr22\_call4.graphlets
file directory: data/chr23\_chr23\_mit.graphlets
file directory: data/chr23\_chr23\_all.graphlets
file directory: data/chr23\_chr23\_rl.graphlets
file directory: data/chr23\_chr23\_call4.graphlets
Number of columns not the same.                picking the minimum value{\ldots}
Number of columns not the same.                picking the minimum value{\ldots}
Number of columns not the same.                picking the minimum value{\ldots}
Number of columns not the same.                picking the minimum value{\ldots}
Number of columns not the same.                picking the minimum value{\ldots}
Number of columns not the same.                picking the minimum value{\ldots}

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}199}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}
          \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
          \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{:}
              \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_19_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}146}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}
          \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{1}
          \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{:}
              \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{n}{count}\PY{p}{)}
              \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
              \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
              \PY{n}{y\PYZus{}error} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
              \PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{yerr}\PY{o}{=}\PY{n}{y\PYZus{}error}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dashed}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{None}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{green}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}title}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_20_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}137}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{73}\PY{p}{)}
          \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
          \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{:}
              \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_21_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}147}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}
          \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{1}
          \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{:}
              \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{n}{count}\PY{p}{)}
              \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
              \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
              \PY{n}{y\PYZus{}error} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
              \PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{yerr}\PY{o}{=}\PY{n}{y\PYZus{}error}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dashed}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{None}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{green}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
              \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}title}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{mit vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{cell})
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_22_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    Compare The cancer cells only:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}151}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}
          \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
          \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{:}
              \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                  \PY{k}{continue}
              \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                  \PY{k}{if} \PY{n}{cell2} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell1} \PY{o}{\PYZgt{}} \PY{n}{cell2}\PY{p}{:}
                      \PY{k}{continue}
                  \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                  \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_24_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}206}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
          \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}
          \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{1}
          \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{:}
              \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                  \PY{k}{continue}
              \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                  \PY{k}{if} \PY{n}{cell2} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell1} \PY{o}{\PYZgt{}} \PY{n}{cell2}\PY{p}{:}
                      \PY{k}{continue}
                  \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{3}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{n}{count}\PY{p}{)}
                  \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                  \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                  \PY{n}{y\PYZus{}error} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                  \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}\PY{p}{)}
                  \PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{yerr}\PY{o}{=}\PY{n}{y\PYZus{}error}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dashed}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{None}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{green}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
                  \PY{n}{ax}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_25_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor} }]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
        \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}
        \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
        \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{:}
            \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                \PY{k}{continue}
            \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                \PY{k}{if} \PY{n}{cell2} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell1} \PY{o}{\PYZgt{}} \PY{n}{cell2}\PY{p}{:}
                    \PY{k}{continue}
                \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{correlations}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{:}\PY{p}{,}\PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}\PY{p}{)}
        \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
        \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    % Add a bibliography block to the postdoc

    \end{document}