### 2018-03-08 papers/temp/transcript.tex

\documentclass{article}
\usepackage[top=1in, bottom=1.25in, left=1in, right=1in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{float}
\usepackage{caption}
\usepackage[font=small, labelfont=bf]{subcaption}
\usepackage{lscape} %landscape
\begin{document}
\doublespacing
\setlength{\parindent}{0pt}
This work is about finding dissimilarities in genome configuration between normal and cancerous cells.

Ideally, we would need to have the 3D structure to make such comparisons, however, such data does not yet exist.

All we have is sets of data that is called HiC data.
Which can be used to come up with matrices of this type that we call contact matrices.

Contact matrices are ...

Let's go into a more detailed discussion...

If you unfold DNA ...

\end{document}