### 2018-05-23 papers/manuscripts/manuscript.tex
final draft of the manuscript

\documentclass[a4,center,fleqn]{NAR}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[super]{nth}
\usepackage{multirow}
% Enter dates of publication
\copyrightyear{2017}
\pubdate{18 March 2017}
\pubyear{2017}
\jvolume{37}
\jissue{12}

\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shadows.blur}
\usetikzlibrary{shapes.symbols}

%\articlesubtype{This is the article type (optional)}

\begin{document}
\title{HiC Contact Map Comaprison Using Graphlet Approach}

\author{%
    Behnam Rasoolian\,$^{1,*}$,
    Debswapna Bhattacharya\,$^{1}$
    \footnote{
    Tel: +1 334 5212814; Email: bzr0014@auburn.edu}}

    \address{%
        $^{1}$ Auburn University
        }
        % Affiliation must include:
        % Department name, institution name, full road and district address,
        % state, Zip or postal code, country

%        \history{%
%            Received January 1, 2009;
%            Revised February 1, 2009;
%            Accepted March 1, 2009}

            \maketitle

\begin{abstract}

In this study, we investigated
dissimilarities between normal cells and leukemic cells
in terms of their three dimensional conformation.
We first thresholed HiC data corresponding to
one normal cell line and four leukemic cells lines.
We then used the thresholded data to extract the 
first 72 orbits for all loci in each contact map.
We then measured pairwise graphlet distances as
well as pairwise graphlet distribution correlations
for each pair of cells and compared them using 
statistical methods.
Our results show that normal-cancer pairs have
significantly higher dinstances from each other
than cancer-cancer pairs.
Also, we concluded that for certain orbits,
cancer-cancer pairs demonstrate higher correlations
than normal-cancer pairs in terms of orbit frequency
distributions.

\end{abstract}

\section{Introduction}
This reasearch is an effort to compare the three-dimensional
structure of cell genomes. To the best of our knowledge
little research has dealt with quantitative comparison
of spatial conformation of chromosomes.
Ideally, it is desirable to compare 3D structures of 
cell in order to make such comparisons.
However, the main challenge in this regard
is that 
3D structure of a cell is not readily available. 
In order to find dissimilarities in the 3D structure of 
chromosomes, we used HiC dataset.
The HiC method, which was developed by 
\cite{lieberman2009comprehensive}, captures interactions between 
chromosomal fragments in kilobase resolution. 
Based on HiC data, a \textit{contact map} or
\textit{interaction frequency (IF) } matrix can be developed 
between \textit{loci} at a desired resolution.
A cell $IF_{ij}$ in an interaction frequency matrix captures 
the number of interaction detected
in HiC dataset between locus $i$ and locus $j$ in the genome.
An interaction matrix can be used to develop both 
inter- and intra-chromosomal interaction matrices.
We have found statistically significant
differences between interaction matrices of
normal cells and cancerous ones.

Recently, HiC data has been mostyly used
in efforts to predict the 3D conformation
of chromosomes using HiC data. 
\cite{noble2011three, 
rousseau2011three, 
hu2013bayesian, 
varoquaux2014statistical, 
trieu2014large, 
zhang20133d, 
lesne20143d, 
bau2011three, 
adhikari2016chromosome3d}.
These efforts, although different in
approach, usually translate 
interaction frequencies in contact maps as an inverse measure of 
distance. For example,  \cite{adhikari2016chromosome3d} developed
a method, Chromosome3D which uses
distance geometry simulated annealing (DGSA) in order to
reconstruct genome structure. These reconstructed 3D 
structures could be used in order to compare normal and malignant
cell lines. However, the main obstacle to doing so is that the
process of reconstruction is computationally
expensive. Almost all reconstruction methods rely on either 
deterministic or heuristic optimization approaches
in order to find the best
confomation that matches contact maps. Also, there is no real 
way of verifying whether the reconstruction results are accurate.
The only method of verifying such results is
Fluorescence in situ hybridizaiton
(FISH) which can capture the actual
3D configuration of chromosomes.
\cite{gasser2002visualizing, lanctot2007dynamic, mateos2009spatially}
However, this method can only be used locally and cannot map
the whole structure of the chromosomes.

Recently, graphlet comparison has emerged
as a novel method for comparing
large networks in order to
find local similarities in them.
Authors of \cite{prvzulj2007biological} 
provide a new measure of protein-protein
interaction (PPI)
network comparison
based on 73 constraints. The method
is used in order to compare two large
PPI networks in order to detect similarities.

In \cite{milenkoviae2008uncovering} the authors
 provide heuristics to compare two nodes in a graph
based on signature vectors, which are 73-dimensional vectors
$\mathbf{s}^T
= [s_0, s_2, ..., s_{72}]$ where $s_i$ denotes the number of nodes in
the network that are part of an orbit $i$. \\
They concluded that proteins with similar surroundings perform
similar functions.

In \cite{milenkovic2010cancer}, the same author investigates 
cancer-causing genes to find similarities in their signatures. After
clustering the genes based on \textit{signature similarity} criteria,
some clusters contain a lot of cancerous genes.
They use 4 different clustering methods with varying parameters to cluster
the proteins. They then predict the cancer-relatedness of a protein 
$i$ using
an enrichment criteria $\frac{k}{|C_i|}$ where $C_i$ is the cluster
where protein $i$ belongs and $k$ is the number of cancer-causing
proteins in $C_i$ and $|C_i|$ is the size of $C_i$.

In this paper, we applied the same approach of graphlet distance
and graphlet distribution comparison on HiC data
in order investigate genome spatial confomations differences
between normal and leukemic cells.

% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
\enlargethispage{-65.1pt}
% **************************************************************

\subsubsection{Notations}
In this paper, matrices and vectors are represented using bold
capital and bold small letters respectively.
Matrix rows and columns are represented by a \textit{dot}
notation. For example, the $i$th row of matrix $M$ is
denoted by $M_{i.}$ and its $j$th column is represented
by $M_{.j}$.

We denote the set of all contact maps in cell line $K$ with 
$\mathbb{C}^K$. If no particular cell line is addressed, the
subscripts are dropped.
Any arbitrary member of $\mathbb{C}$ is denoted by 
$\mathbf{C}_{ij}$, where $i$ and $j$ ($j \ge i$) 
represent the two chromosomes involved. 
In human cells this set contains a total of 276 contact maps,
23 of which are intra-chromosomal and the rest are inter-chromosomal.
For ease of representations, intra-chromosomal contact maps are
distinguished by a single superscript, so we have $\mathbf{C}_{i,i} =
\mathbf{C}_i$.

We denote the number of loci in a chromosome $i$ by $N_i$.
The set of all loci involved in contact map $C_{ij}$ is denoted 
by $\mathbb{V}_{ij}$.
In intra-chromosomal contact maps, $\mathbb{V}_{i,i}$ containts only the 
loci of that particular chromosome ($|\mathbb{V}_i| = N_i$), while in 
inter-chromosomal contact maps $\mathbb{V}_{ij}$ contains the loci in
the both of chromosomes involved 
($\mathbb{V}_{ij} = \mathbb{V}_i \cup \mathbb{V}_j$).

\section{Materials and Methods}
We re-used Leukemic Hi-C libraries 
created in \cite{wang2013properties}
These libraries were sequenced 
for cases of primary human B-acute
lymphoblastic leukemia (B-ALL or ALL), 
the MHH-CALL-4 B-ALL cell
line (CALL4), 
and the follicular lymphoma cell-line (RL).
As in \cite{wang2013properties}, 
We used normal B-cell line (GM068990)
from \cite{lieberman2009comprehensive} 
for our comparisons.
We created contact maps of resolution
500kb and normalized it using 
the \texttt{iced} package in python
developed by \cite{servant2015hic}.
We also performed and extra normalization by
performing network deconvolution \cite{feizi2013network}
in order to remove
the effect of indirect interactions.

\subsection{Thresholding contact maps}
In order to be able to extract graphlets, HiC contact maps should be modeled as
unweighted graphs where the nodes represent the loci and an edge between two 
nodes represent a \textit{significant} interaction between the loci.
This can be achieved by thresholding the contact maps. The result
of the thresholding procedure is a binary matrix which also can serve as
an adjacency matrix for an unweighted, undirected graph. The graph can then be
used for orbit extraction.

When thresholding contact maps, it is necessary to make sure
that both global and local features are maintained. We could consider 
thresholding the contact maps by simply setting values above a fixed value to
one and the rest to zero; However, in practice, this method resulted in graphs
that capture the local structure of the contact maps poorly. This is because
intensities follow an exponential distribution with a mean close to zero
with a few very larges values that correspond to interactions along 
or close to the main diagonal of the contact maps.
Thus, picking relatively large numbers would result in ignoring interactions
that are far from the main diagonal while picking small values will lead to
capturing too many (insignificant) interactions.

To the best of our knowledge, little work has dealt with the task of
thresholding HiC contact maps. There has been some statistical approaches
developed on similar data in other fields. 
For example, authors of \cite{ginestet2011statistical} developed 
Statistical Network (SPN) analysis
where the choice of thresholding value is made by statistical inference.
This method, although very robust, works within the framework of 
design of experiments where the same
network can be extracted for different 
individuals under different treatments. Thus a relatively large
set of different contact maps need to be available in order for
this method to be applicable towards our end.

Instead, in order to threshold the matrix so that both global and local patterns are
captured, we borrowed the concept of \textit{adaptive thresholding} from image 
processing context. In this method, in order to be set, a pixel should have
an intensity larger than the average of non-zero intensities in its
\textit{neighborhood}. The neighborhood is defined by a sliding kernel 
that passes through the contact map with the pixel at its middle at 
each step. Figure \ref{local_thresholded_chr1_chr1} demonstrates result of 
this thresholding approach for intra-chromosomal contact maps of chromosome 1.
Refer to supplementary material for all 23 interchromosomal thresholding
results.
\begin{figure}[t]
    \centering
    \includegraphics[width=.47\textwidth]{figures/local_thresholded_chr1_chr1.png}
    \caption{Result of thresholding interchromosomal contact map of chromosome 1
    using a kernels of size $5 \times 5$ for all cell lines. 
    The first row shows the thresholded
    maps. Second and third rows demonstrate pair-wise similarities and 
    differences between contact maps respectively.}
    \label{local_thresholded_chr1_chr1}
\end{figure}

\subsection{Orbit Extraction}
Once the thresholded contact maps are obtained, graphlets and orbits can be 
extracted. We used the \texttt{orca} package in \texttt{R} programming 
language to extract the graphlets. As a result of graphlet extraction, 
For each loci in each contact map, a \textit{signature vector} of size
73 is created. Thus for each cell line, we would have 276 
\textit{signature matrices} of
size $|V^{ij}|\times 73$, where $V^{ij}$ is the number of loci
involved in contact map between chromosomes $i$ and $j$. Figure
\ref{fig:graphlet_extraction} illustrates the process and results
of signature matrix extraction schematically.
\begin{figure}
    \centering
    \includegraphics[width=.47\textwidth]{figures/graphlet_extraction.png}
    \caption{Graphlet extraction for the four cell lines. For each
    loci in each contact map between chromosomes $i$ and $j$, 
    the signature vectors of length 73 are extracted, resulting 
    in a \textit{signature matrix} of size $|V^{ij}| \times 73$,
    where $V^{ij}$ 
    is the number of loci involved.}
    \label{fig:graphlet_extraction}
\end{figure}

For a particular
$\mathbf{C}_{ij}$, we denote $\mathbf{S}_{ij}$ as its \textit{signature 
matrix}. Each cell $S_{ijlo}$ in $\mathbf{S}_{ij}$ captures how many
times loci $l$ in $\mathbf{C}_{ij}$ occured as part of orbit $o$.

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figures/graphlet_distance_schema.png}
    \caption{Calculating pair-wise loci distances. For each loci (row) 
    in each
    contact map in MIT cell line, its distance is calculated based on
    equation \ref{eq:distance_total} with the corresponding loci in
    leukemic cells. The result of this process is a 
    \textit{signature distance vector} of size
    $|V^{ij}| = N_i+N_j$ for each contact map.
    }
    \label{graphlet_distance_schema}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figures/graphlet_correlation_schema.png}
    \caption{Calulating pair-wise orbit correlations. For each orbit (column)
    in each contact map in MIT cell line, its correlation with the
    same orbit in the same contact map 
    in leukemic cells is calculated. The result of
    this process is a \textit{signature correlation} vector of size
    73 which captures how similar frequencies of two orbits are.}
    \label{graphlet_correlation_schema}
\end{figure}
%$
%\begin{pmatrix}
%    {S}^A_{\scriptscriptstyle i,j,1,1}         &
%    \cdots                                     &    
%    {S}^A_{\scriptscriptstyle i,j,1,72}        \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    \vdots                                     &                
%                                               &      
%    \vdots                                     \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   {S}^A_{\scriptscriptstyle i,j,N_1,1}    &
%   \ddots                                            &    
%   {S}^A_{\scriptscriptstyle i,j,N_1, 72} \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    \vdots                                     &                
%                                               &      
%    \vdots                                     \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   {S}^A_{\scriptscriptstyle i,j,N_1+N_2,1}   &
%    \cdots                                &
%   {S}^A_{\scriptscriptstyle i,j,N_1+N_2, 72}
%\end{pmatrix}
%$
%
%$
%\begin{pmatrix}
%    {C}^A_{\scriptscriptstyle i,j,1,1}     &
%    \cdots    &
%    {C}^A_{\scriptscriptstyle i,j,1,N_1} \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    \vdots                             &
%    \ddots&
%    \vdots                             \\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    {C}^A_{\scriptscriptstyle i,j,N_1,1}   &
%    \cdots&                  
%    S^A_{\scriptscriptstyle i,j,N_1, N_1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\end{pmatrix}
%$

%We divided the task of graphlet comparison into two parts: first we compare
%graphlets from each contact map in normal cell lines (MIT) with the 
%same contact maps in the other three Leukemic cells. Second we compare 
%contact maps in a similar way but this time only between leukemic cells.
%In the former case, the null hypothesis is that there is no difference
%between contact maps of normal cells and leukemic cells and in the latter
%case, the null hypotheis is that there is no difference between 
%different leukemic cells.

We consider two measures of \textit{difference} when comparing contact
map graphlets across cell lines. 
The first measure is \textit{signature distance vectors} between
each contact map of two cell lines. 
For a pair of cells A and B, let 
$\mathbf{S}^A_{ij}$  and $\mathbf{S}^B_{ij}$ be their
signature matrices. The \textit{signature distance} of
contact map $\mathbf{C}_{i,j}$ between A and B is denoted
by $\mathbf{d}^{\scriptscriptstyle A,B}_{ij}$. $\mathbf{d}^{\scriptscriptstyle A,B}_{ij}$ 
is a vector of size $|V_{i,j}|$
and its elements $d^{\scriptscriptstyle A,B}_{i,j,l}$ are
calculated using the following formula from \cite{prvzulj2007biological}:

\begin{equation}
    d^{\scriptscriptstyle A,B}_{i,j,l} = 
    \frac{1}{73}\sqrt{\sum_{o=0}^{72}{t_{lo}^2}}
    \label{eq:distance_total}
\end{equation}

where elements of $t_{i,j,l,o}$ is the
distance between each
loci (row) $l$ in $\mathbf{S}^A$ and the the same loci in 
$\mathbf{S}^B$ for
orbit $o$ as is calculated as below:

\begin{equation}
    t_{lo} = w_o \times 
    \frac{log(S_{ijlo}^A+1) - log(S_{ijlo}^B+1)}
    {log(max(S_{ijlo}^A, S_{ijlo}^B) + 2)}
    \label{eq:distance_single}
\end{equation}

This process is illustrated in Figure \ref{graphlet_distance_schema}.
Using this distance measure, we can quantify how two loci are close to
each other in terms of local neighborhood between the two contact maps.

The second measure of comparison that we use captures how 
similar two orbits are in terms of their count 
frequencies across loci between two contact maps. 
Each column in $S_{ij}$ can provide information
regarding the \textit{frequency distribution} of orbits throughout
the contact map $C_{ij}$. 
We can find how similar these distributions are to each other using
correlation measures.
These correlations are denoted by $\mathbf{m}^{\scriptscriptstyle A,B}_{i,j}$ and
can be calculate using
any plausible correlation measure. 
In this study, for each contact map,  we calculated
similarity between orbit distributions using Pearson's r 
correlation, which is computationally efficient.
However, pearson's r might not be able to capture
non-functional relationships between distributions. As a result, we
also used Maximal Information Coefficient (MIC) 
\cite{reshef2011detecting} in order to compare
correlations. MIC calculates mutual information (MI) between two
distributions, but utilizes dynamic programming in order adjust
bin sizes and numbers in order to achieve highest MI.
MIC values between two variables fall between 0 and 1,
with 0 meaning the two variables are completely independent
and 1 meaning one is dependant on the other.
We used both Pearson's r and MIC in order to compare orbit
frequencies. Although results from both approaches were more
or less consistent, MIC showed higher robustness than Pearson's 
r method.

If MIC is used as correlation measure, each element of 
 $\mathbf{c}$ is calculated as below:
\begin{equation}
    m^{\scriptscriptstyle A,B}_{ijo} = MIC(\mathbf{S}^A_{ij.o}, \mathbf{S}^B_{ij.o})
    \label{eq:mic}
\end{equation}
Alternatively, if we use Pearson criterion we would have:
\begin{equation}
    m^{\scriptscriptstyle A,B}_{ijo} = Pearson(\mathbf{S}^A_{ij.o}, \mathbf{S}^B_{ij.o})
    \label{eq:pearson}
\end{equation}

\section{results and discussions}
\subsection{Contact Map Orbit Vector Distance}
\begin{figure}
    \centering
    \begin{subfigure}[b]{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/orbit_distances_normal_vs_cancer.png}
        \caption{}
        \label{fig:orbit_distances_all_a}
    \end{subfigure}

    \begin{subfigure}[b]{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/orbit_distances_cells.png}
        \caption{}
        \label{fig:orbit_distances_all_b}
    \end{subfigure}

    \caption{   
        \textbf{Average Pair-wise graphlet signature difference 
        over all 276 contact maps:}
        Each point on a graph is the result of averaging all
        the distances across all loci of all contact maps for 
        each pair of cells. A 0.1 standard deviation error
        bar is also plotted for each point.
        ($\bar{\mathbf{d}}^{\scriptscriptstyle A,B}$).  
     }
    \label{fig:orbit_distances_all}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}[b]{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/orbit-distances_chr14.png}
    \end{subfigure}
    \caption{   
        Pair-wise graphlet signature distances for all contact maps of
        chromosome 14. Warm colors as used for cancer-cancer pairs and
        cold colors are used for normal-cancer pairs. As can be seen
        cancer-cancer pairs tend to be close to each other than to the
        normal cell. This is specially true for the first 13 
        inter-chromosomal contact maps.
     }
    \label{fig:orbit_distances_chr14}
\end{figure}

By comparing signature distance vectors, one can find
how contact maps
differ from each other in terms of local structure. 
Contact maps can serve as measures of spatial proximity between
loci. Graphlets capture certain patterns of interaction, or
in other words, spatial neighborhood for each loci. Thus, if
signature vectors of two loci are close,
 it can be inferred that they have 
similar spatial neighborhood.

We can compare pairs of contact maps in terms of their closeness
to each other. 
As an example, in figure
\ref{fig:orbit_distances_chr14}, all pairs of cells are
compared to each other in terms of their distance for contact
maps involving chromosome 14. We can se that 
for the first 13 inter-chromosomal 
contact maps, ALL and RL cell lines are closer to each other
than to other cell lines.

We performed one-way ANOVA statistical test to 
see if there are significant 
difference between cancer-normal and cancer-cancer pairs.
We found that the difference statistically significant difference.
($F(1, 1654) = 20.49, p < 0.0001$).
As illustrated in figure \ref{fig:orbit_distances_all_a},
we can see that normal-cancer pairs have higher distance
from each other than cancer-cancer pairs.

We then continued to investigate each pair separately to
see if there is any significant difference between them.
Again, our statistical tests (ANOVA) showed significant difference
between pairs of cells.
The results are shown in figure \ref{fig:orbit_distances_all_b}.
We can see that ALL-RL pair are closest to each other while
ALL-MIT and MIT-RL are most distant. We found statistcally
significant difference for difference between individual
pairs except for ALL-CALL4 and CALL4-RL as well as
MIT-RL and ALL-MIT. The results of thse tests 
can be found in supplementary material.

The results in figure \ref{fig:orbit_distances_all_b} is also
in keeping with what we see in figure \ref{fig:orbit_distances_chr14}. For example, as mentioned earlier, our results show that
on average, ALL
and RL cell lines are closer to each other than to other
cell lines, which is also the case in figure
\ref{fig:orbit_distances_chr14} for majority of contact maps.

\subsection{MIC Comparison}

\begin{figure*}
    \centering
%    \begin{subfigure}[b]{\textwidth}
%        \includegraphics[width=\textwidth]{figures/contact_maps_correlations_cancer_vs_normal.png}
%    \caption{}
%    \label{fig:contact_map_correlations_all}
%    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{figures/orbits_correlations_cancer_vs_normal.png}
    \caption{}
    \label{fig:orbits_correlations_normal_vs_cancer}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{figures/orbits_correlations_all.png}
    \caption{}
    \label{fig:orbits_correlations_all}
    \end{subfigure}
    \caption{
        \textbf{Pair-wise 
        average contact map orbit correlations
        for all contact maps:}
        ($\bar{\mathbf{m}}^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j} 
        \quad \forall i,j \in \{1 ... 23\} \quad \& \quad j \ge i$:
        average along the 
        red \textit{vertical} arrow in figure 
        \ref{graphlet_correlation_schema}).
        These values are calculated by averaging over 
        pairwise correlations of orbits of
        $\mathbb{Q}$ in a contact map.
        \textbf{Pair-wise average orbit correlations:}
        In figure \ref{fig:orbits_correlations_all}, each point
        in the graph is the result of averaging pair-wise
        orbit correlations over all contact maps
        ($\frac{1}{276}\sum_{i=0}^{23}\sum_{j=i}^{23}{
        m^{\scriptscriptstyle \scriptscriptstyle A,B}_{i,j, o}} \quad 
        \forall o \in \{0, 1, ..., 72\}$:
        average along the red \textit{horizontal} arrows in figure 
        \ref{graphlet_correlation_schema}).
        Counts for certain orbits are always zero in inter-chromosomal
        maps, leading to average value close to zero in 
        Figure \ref{fig:orbits_correlations_all}.
        In this figure, cancer cells data points are depicted
        in warm colors while normal cells are depicted in
        cold colors in increased contrast. As can be seen
        orbit distributions are more similar to each other
        for cancer cells.
    }
    \label{fig:orbits_correlations}
\end{figure*}

In addition to comparing cells in terms of their
orbit distances, we can compare them by measuring how
often certain graphlets occur in their contact maps. By
doing so, we measure the frequency distribution of the spatial
structures represented by orbits in each contact map. In order
to see how closely such structures are distributed, we can
compare contact maps by calculating the correlation between
their orbit distributions. A higher correlation for certain orbits
would mean higher similarity in terms of that particular spatial
structure between the loci involved.

Before going on with the results,
it is worth mentioning that interchromosomal 
thresholded contact maps 
represent
a bipartide graph with the loci from each chromosome 
on one side. Due to this
bipartide nature of the graphs in inter-chromosomal maps,
count of certain orbits is always 0, resulting in
a correlation values of 0 for them as well.
You can see the bias in 
figure \ref{fig:orbits_correlations} where 
average correlations of orbits
$\mathbb{Q} = \{3, 9, 10{\text -}14, 20{\text -}34, 39{\text -}48, 51{\text -}72\}$ 
are close to zero. In fact all correlations
corresponding to these orbits are 0 except for the ones between 
the same 
chromosomes.

We caclulated pair-wise MIC values 
for each orbit in each of the 276
contact maps from MIT, ALL, RL, and CALL4 data separately. 
We found statistically significant difference between
cancer-cancer and normal-cancer correlations.
($F(73, 1582)=6.29$, $p < 0.00001$, Wilk's $\Lambda = 0.775$)
 The difference  is also illustrated in figure
\ref{fig:orbits_correlations_normal_vs_cancer},
which plots corrected
mean difference of MIC values for cance-cancer and 
normal-cancer correlations.  
We performed statistical test to see if there is a 
significant difference correlation between individual pairs
of cells. Figure \ref{fig:orbits_correlations_all} demonstrate
such difference.
Average correlation over all contac maps
for normal-cancer pairs are smaller
than cancer-cancer pairs. This is corroborated
by the results of statical test which verify that such difference
is in fact signifcant.($F(365, 7882) = 3.91$, $p < 0.00001$, 
Wilk's $\Lambda = 0456$)
Figures \ref{fig:orbits_correlations_normal_vs_cancer} and
\ref{fig:orbits_correlations_all} both show more details about
this difference in correlation. As can be observed, for obits
in $\mathbb{Q}$, normal correlations are smaller than cancer
correlations, while for the rest of the orbits, there is no
difference.

Figure \ref{fig:orbits_correlations} demonstrates that certain
orbits of Leukemic cells
have higher correlation to each other 
than to the
normal MIT cell. In fact our statistical analysis shows that 
\textit {for orbits NOT in $\mathbb{Q}$, 
intra-leukemic orbit correlations are significantly higher
than leukemic-normal orbit correlations}. This implies
there are significant differences between normal and
leukemic cells in terms of their local structure.

\section{Resources}
\textbf{Hi-C Datasets:}
\begin{enumerate}
    \item \href{https://github.com/rasoolianbehnam/watson}{Code base for this article}
    \item \href{http://sysbio.rnet.missouri.edu/T0510/tmp_download/link_to_download_genome_data/}
        {Datasets including cancerous cells}
    \item \href{https://bcm.app.box.com/v/aidenlab/folder/11234760671}{Original Datasets}
\end{enumerate}

\bibliography{lit}
\bibliographystyle{unsrt}
\end{document}