### 2018-07-11 papers/manuscripts/supplementary/02_from_graphlets_to_analysis.tex

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{02\_from\_graphlets\_to\_analysis}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{Local Thresholding}\label{local-thresholding}

Here, We extract orbits for a signle chromosome in all 4 cell
lines, using a different method of thresholding. In order to do this, we
slide a kernel of size k through each pixel and check if it has a
particular property. If the property is satisified, then the pixel is
set to 1, otherwise it is set to 0. The two properties that I experiment
with are \emph{maximum thresholding} and \emph{normal thresholding}. 1.
\textbf{Maximum Thresholding}: in max thresholding, if the pixel is the
maximum of its neighbors with respect to the kernel, then it is set.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  \textbf{Normal Thresholding}: in normal thresholding, if the pixel is
  larger that the average of the neighbors then it is set.
\end{enumerate}

    \section{Loading Data}\label{loading-data}

The first step is to load the raw Hi-C contact maps. We have already
extracted all inter- and intra-chromosomal contact maps and compiled
them into as single large numpy array format.

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_3_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \section{Selecting a inter-/intra-chromosomal contact
map}\label{selecting-a-inter-intra-chromosomal-contact-map}

You can chang chr1 and chr2 variables based on the contact map you want
to investigate.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{1}
        \PY{n}{chr2} \PY{o}{=} \PY{l+m+mi}{2}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}          \PY{o}{=} \PY{n}{get\PYZus{}contact\PYZus{}map}\PY{p}{(}\PY{n}{mit\PYZus{}full}        \PY{p}{,} \PY{n}{lengths\PYZus{}low\PYZus{}res}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}          \PY{o}{=} \PY{n}{get\PYZus{}contact\PYZus{}map}\PY{p}{(}\PY{n}{all\PYZus{}full}        \PY{p}{,} \PY{n}{lengths\PYZus{}low\PYZus{}res}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}           \PY{o}{=} \PY{n}{get\PYZus{}contact\PYZus{}map}\PY{p}{(}\PY{n}{rl\PYZus{}full}         \PY{p}{,} \PY{n}{lengths\PYZus{}low\PYZus{}res}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}        \PY{o}{=} \PY{n}{get\PYZus{}contact\PYZus{}map}\PY{p}{(}\PY{n}{call4\PYZus{}full}      \PY{p}{,} \PY{n}{lengths\PYZus{}low\PYZus{}res}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
        \PY{n}{showImages}\PY{p}{(}\PY{n}{data}\PY{p}{,} \PY{n}{rows} \PY{o}{=} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{titles}\PY{o}{=}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of rows and columns: 2, 2
(4, 4)

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_5_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsection{Cleaning Data}\label{cleaning-data}

As can be observed above, there are several rows and columns that simply
don't contain any data and are all zeros, in the following code we clean
the matrices to remove these zero rows and columns since they will cause
problems in future analysis.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{c+c1}{\PYZsh{}There are some blank rows and columns in the matrix. let\PYZsq{}s remove them.}
        \PY{n}{blankRows0}\PY{p}{,} \PY{n}{blankCols0} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankRows1}\PY{p}{,} \PY{n}{blankCols1} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankRows2}\PY{p}{,} \PY{n}{blankCols2} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankRows3}\PY{p}{,} \PY{n}{blankCols3} \PY{o}{=} \PY{n}{getBlankRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankRows} \PY{o}{=} \PY{n}{Set}\PY{p}{(}\PY{p}{[}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankCols} \PY{o}{=} \PY{n}{Set}\PY{p}{(}\PY{p}{[}\PY{p}{]}\PY{p}{)}
        \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows0}\PY{p}{)}
        \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows1}\PY{p}{)}
        \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows2}\PY{p}{)}
        \PY{n}{blankRows}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankRows3}\PY{p}{)}
        \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols0}\PY{p}{)}
        \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols1}\PY{p}{)}
        \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols2}\PY{p}{)}
        \PY{n}{blankCols}\PY{o}{.}\PY{n}{update}\PY{p}{(}\PY{n}{blankCols3}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
        \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{removeRowsAndColumns}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{blankRows}\PY{p}{,} \PY{n}{blankCols}\PY{p}{)}
        \PY{n}{n1}\PY{p}{,} \PY{n}{m1} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
        \PY{n}{n2}\PY{p}{,} \PY{n}{m2} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
        \PY{n}{n3}\PY{p}{,} \PY{n}{m3} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{ALL}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
        \PY{n}{n4}\PY{p}{,} \PY{n}{m4} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
('size of old matrix:', (495, 486))
('size of new matrix:', (454, 479))
('size of old matrix:', (495, 486))
('size of new matrix:', (454, 479))
('size of old matrix:', (495, 486))
('size of new matrix:', (454, 479))
('size of old matrix:', (495, 486))
('size of new matrix:', (454, 479))

    \end{Verbatim}

    \subsubsection{Comparing CALL4 with RL low-resolution
data:}\label{comparing-call4-with-rl-low-resolution-data}

Feel free to change D1 and D2 to pick any data from set of ```{[}'MIT',
'ALL', 'RL', 'CALL4'{]}````.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}6}]:} \PY{n}{D1} \PY{o}{=} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{CALL4}\PY{l+s+s1}{\PYZsq{}}
        \PY{n}{D2} \PY{o}{=} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{RL}\PY{l+s+s1}{\PYZsq{}}
        \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{20}\PY{p}{)}
        \PY{c+c1}{\PYZsh{} Size of the kernel}
        \PY{n}{k} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}
        \PY{c+c1}{\PYZsh{} can be either \PYZsq{}max\PYZsq{} for setting the maximum value in each kernel to 1 and the rest to 0}
        \PY{c+c1}{\PYZsh{} or \PYZsq{}normal\PYZsq{} for setting all values above mean + t * std withing the kernel to 1 and }
        \PY{c+c1}{\PYZsh{} the rest to 0}
        \PY{n}{method} \PY{o}{=} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal}\PY{l+s+s1}{\PYZsq{}}
        \PY{c+c1}{\PYZsh{} in case of normal thresholding, t is the coefficient }
        \PY{c+c1}{\PYZsh{} of the standard deviation; that is, n each kernel iteration K}
        \PY{c+c1}{\PYZsh{} if K[i, j] \PYZgt{} mean(K) + t * std(K), then it is set to 1.}
        \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{0}
        \PY{n}{params} \PY{o}{=} \PY{n+nb+bp}{None}
        \PY{n}{symmetric}\PY{o}{=} \PY{n}{chr1}\PY{o}{==}\PY{n}{chr2}
        \PY{n}{D1\PYZus{}os} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{p}{(}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{D1}\PY{p}{]}\PY{o}{+}\PY{l+m+mi}{1}\PY{o}{+}\PY{l+m+mf}{1e\PYZhy{}5}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{n}{k}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{n}{method}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{n}{t}\PY{p}{,} \PY{n}{params}\PY{o}{=}\PY{n}{params}\PY{p}{,} \PY{n}{symmetric}\PY{o}{=}\PY{n}{symmetric}\PY{p}{)}
        \PY{n}{D2\PYZus{}os} \PY{o}{=} \PY{n}{local\PYZus{}threshold}\PY{p}{(}\PY{p}{(}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{D2}\PY{p}{]}\PY{o}{+}\PY{l+m+mi}{1}\PY{o}{+}\PY{l+m+mf}{1e\PYZhy{}5}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{k} \PY{o}{=} \PY{n}{k}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{n}{method}\PY{p}{,} \PY{n}{t} \PY{o}{=} \PY{n}{t}\PY{p}{,} \PY{n}{params}\PY{o}{=}\PY{n}{params}\PY{p}{,} \PY{n}{symmetric}\PY{o}{=}\PY{n}{symmetric}\PY{p}{)}
        \PY{n}{n1}\PY{p}{,} \PY{n}{m1} \PY{o}{=} \PY{n}{D1\PYZus{}os}\PY{o}{.}\PY{n}{shape}
        \PY{n}{n2}\PY{p}{,} \PY{n}{m2} \PY{o}{=} \PY{n}{D2\PYZus{}os}\PY{o}{.}\PY{n}{shape}
        \PY{n}{n} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{[}\PY{n}{n1}\PY{p}{,} \PY{n}{n2}\PY{p}{]}\PY{p}{)}
        \PY{n}{m} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{[}\PY{n}{m1}\PY{p}{,} \PY{n}{m2}\PY{p}{]}\PY{p}{)}
        \PY{n}{D1\PYZus{}os} \PY{o}{=} \PY{n}{D1\PYZus{}os}\PY{p}{[}\PY{p}{:}\PY{n}{n}\PY{p}{,} \PY{p}{:}\PY{n}{m}\PY{p}{]}
        \PY{n}{D1\PYZus{}data} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{n}{D1}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{n}{n}\PY{p}{,} \PY{p}{:}\PY{n}{m}\PY{p}{]}
        \PY{n}{D2\PYZus{}os} \PY{o}{=} \PY{n}{D2\PYZus{}os}\PY{p}{[}\PY{p}{:}\PY{n}{n}\PY{p}{,} \PY{p}{:}\PY{n}{m}\PY{p}{]}
        \PY{n}{D2\PYZus{}data} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{n}{D2}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{n}{n}\PY{p}{,} \PY{p}{:}\PY{n}{m}\PY{p}{]}
        \PY{n}{a} \PY{o}{=} \PY{p}{(}\PY{p}{(}\PY{n}{D1\PYZus{}os} \PY{o}{*} \PY{n}{D1\PYZus{}data}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{l+m+mi}{1}
        \PY{n}{b} \PY{o}{=} \PY{p}{(}\PY{p}{(}\PY{n}{D2\PYZus{}os} \PY{o}{*} \PY{n}{D2\PYZus{}data}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{l+m+mi}{1}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(2, 2, 2, 2)
(2, 2, 2, 2)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}8}]:} \PY{n}{images} \PY{o}{=} \PY{p}{[} \PY{n}{D1\PYZus{}os}\PY{p}{,} \PY{n}{a}\PY{p}{,} \PY{p}{(}\PY{n}{a}\PY{o}{\PYZhy{}}\PY{n}{b}\PY{p}{)} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{D2\PYZus{}os}\PY{p}{,}  \PY{n}{b}\PY{p}{,} \PY{n}{a} \PY{o}{*} \PY{n}{b} \PY{p}{]}
        \PY{n}{titles} \PY{o}{=} \PY{p}{[}\PY{n}{D1}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Thresholded }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s1}{\PYZsq{}}\PY{o}{\PYZpc{}}\PY{k}{D1}, \PYZsq{}Difference\PYZsq{}, D2, \PYZsq{}Thresholded \PYZpc{}s\PYZsq{}\PYZpc{}D2, \PYZsq{}Similarity\PYZsq{}]
        \PY{n}{showImages}\PY{p}{(}\PY{n}{images}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{titles}\PY{o}{=}\PY{n}{titles}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of rows and columns: 2, 3
(6, 6)

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_10_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    % Add a bibliography block to the postdoc

    \section{Graphlet Correlation
Analysis}\label{graphlet-correlation-analysis}

Remember, in notebook \emph{Thresholding}, I thresholded contact maps
and saved so that orca can read them.

In this notebook, I perform two actions: 1. I apply \texttt{count5}
procedure in orca to extract 73 orbits for each loci in each chromosome.
After this stage, graphlets will be save with \texttt{.graphlets}
extension in the \texttt{./data} folder.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  I then read the graphlets and find both loci-wise distance and
  graphlet-wise correlations between the four cell lines.
\end{enumerate}

By repeating steps 1 and 2 for all 23 chromosomes, I will have pairwise
MIC values for all orbits of all cell lines. I would eventually be able
to have a data array of shape \texttt{A(23\ *\ 73)}, for each pair of
cell lines where \texttt{A{[}i,\ j{]}} denotes the correlation between
the jth orbit in the ith chromosome.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}1}]:} \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k+kn}{as} \PY{n+nn}{np}
        \PY{k+kn}{import} \PY{n+nn}{cv2}
        \PY{k+kn}{from} \PY{n+nn}{library.utility} \PY{k+kn}{import} \PY{o}{*}
        \PY{k+kn}{import} \PY{n+nn}{matplotlib.pyplot} \PY{k+kn}{as} \PY{n+nn}{plt}
        \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n}{normalization}
        \PY{k+kn}{from} \PY{n+nn}{iced} \PY{k+kn}{import} \PY{n+nb}{filter}
        \PY{k+kn}{import} \PY{n+nn}{subprocess}
        \PY{k+kn}{from} \PY{n+nn}{scipy.interpolate} \PY{k+kn}{import} \PY{n}{spline}
        \PY{o}{\PYZpc{}}\PY{k}{load\PYZus{}ext} autoreload
        \PY{o}{\PYZpc{}}\PY{k}{autoreload} 2
        \PY{o}{\PYZpc{}}\PY{k}{pylab} inline
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Populating the interactive namespace from numpy and matplotlib

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/local/lib/python2.7/site-packages/IPython/core/magics/pylab.py:161: UserWarning: pylab import has clobbered these variables: ['colors', 'log']
`\%matplotlib` prevents importing * from pylab and numpy
  "\textbackslash{}n`\%matplotlib` prevents importing * from pylab and numpy"

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}2}]:} \PY{n}{chr2count} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{n}{count2chr} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
        \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
            \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]} \PY{o}{=} \PY{n}{count}
                \PY{n}{count2chr}\PY{p}{[}\PY{n}{count}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
                \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}

        \PY{n}{sameIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
        \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
            \PY{n}{sameIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
        \PY{c+c1}{\PYZsh{}print(\PYZdq{}Same indices: \PYZpc{}s\PYZbs{}n\PYZdq{}\PYZpc{}sameIndices)}
        \PY{n}{differentIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
        \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
            \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{n}{differentIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
        \PY{c+c1}{\PYZsh{}print(\PYZdq{}Different indices: \PYZpc{}s\PYZdq{}\PYZpc{}differentIndices)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{data\PYZus{}names} \PY{o}{=} \PY{p}{[}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{MIT}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ALL}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{RL}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{CALL4}\PY{l+s+s2}{\PYZdq{}}\PY{p}{]}
        \PY{n}{root} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{..}\PY{l+s+s2}{\PYZdq{}}
        \PY{n}{file\PYZus{}dir\PYZus{}template} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{/extracted\PYZus{}all\PYZus{}contact\PYZus{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{root}\PY{p}{,}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{file\PYZus{}name\PYZus{}template} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{/all\PYZus{}in\PYZus{}one\PYZus{}500kb.npy}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{file\PYZus{}dir\PYZus{}template}\PY{p}{)}
        \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{File Directory: }\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{file\PYZus{}dir\PYZus{}template})
        \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{File Name Template: }\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+se}{\PYZbs{}\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{k}{file\PYZus{}name\PYZus{}template})

        \PY{n}{mit\PYZus{}colors} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}0d75f8}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}42b395}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}99cc04}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
        \PY{n}{cancer\PYZus{}colors} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}ab1239}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}c83cb9}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{\PYZsh{}cf524e}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
File Directory: "../extracted\_all\_contact\_\%s"
File Name Template: "../extracted\_all\_contact\_\%s/all\_in\_one\_500kb.npy"

    \end{Verbatim}

    \subsection{Extracting orbits from thresholded
files}\label{extracting-orbits-from-thresholded-files}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{c+c1}{\PYZsh{} Uncommenct the following code to extract graphlets from contact maps}
        \PY{c+c1}{\PYZsh{}for chr1 in range(1, 23):}
        \PY{c+c1}{\PYZsh{}    for chr2 in range(chr1, 23):}
        \PY{c+c1}{\PYZsh{}        command = \PYZsq{}Rscript ../external/rscript\PYZus{}intra.r \PYZpc{}s \PYZpc{}s\PYZsq{}\PYZpc{}(chr1, chr2)}
        \PY{c+c1}{\PYZsh{}        output = subprocess.check\PYZus{}output(command, subprocess.STDOUT, shell=True)}
        \PY{c+c1}{\PYZsh{}        print(output)}
\end{Verbatim}

    \subsection{Reading Already saved
orbits}\label{reading-already-saved-orbits}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{n}{graphlets} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
            \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
            \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
            \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
                \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                    \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
                    \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{/data/graphlets/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{root}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{,} \PY{n}{cell}\PY{o}{.}\PY{n}{lower}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} 
                    \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
\end{Verbatim}

    \subsection{Putting extracted graphlets in
order}\label{putting-extracted-graphlets-in-order}

The following code block assembles all 23 orbit signature vectors of a
loci into a \(23 \times 73\) matrix.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}6}]:} \PY{n}{lengths} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{k}{for} \PY{n}{key} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
            \PY{k}{print}\PY{p}{(}\PY{n}{key}\PY{p}{)}
            \PY{n}{lengths}\PY{p}{[}\PY{n}{key}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{load}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../numpy\PYZus{}data/length\PYZus{}low\PYZus{}res.npy}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}

        \PY{n}{num\PYZus{}loci} \PY{o}{=} \PY{l+m+mi}{100000}
        \PY{n}{loci\PYZus{}data} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
            \PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
            \PY{n}{num\PYZus{}loci} \PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{[}\PY{n}{np}\PY{o}{.}\PY{n}{sum}\PY{p}{(}\PY{n}{lengths}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{num\PYZus{}loci}\PY{p}{]}\PY{p}{)}\PY{p}{)}
            \PY{k}{print} \PY{n}{num\PYZus{}loci}
            \PY{k}{for} \PY{n}{l} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{num\PYZus{}loci}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{:}
                \PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{l}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{l+m+mi}{24}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}

        \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
            \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                    \PY{n}{l1} \PY{o}{=} \PY{n}{lengths}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{,} \PY{p}{:}\PY{p}{]}
                    \PY{n}{l2} \PY{o}{=} \PY{n}{lengths}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{,} \PY{p}{:}\PY{p}{]}
                    \PY{n}{temp} \PY{o}{=} \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]}
                    \PY{k}{if} \PY{n}{chr1} \PY{o}{==} \PY{n}{chr2}\PY{p}{:}
                        \PY{k}{for} \PY{n}{l} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{temp}\PY{p}{)}\PY{p}{)}\PY{p}{:}
                            \PY{n}{loci\PYZus{}abs\PYZus{}num} \PY{o}{=} \PY{n}{l1}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]} \PY{o}{+} \PY{n}{l}
                            \PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{loci\PYZus{}abs\PYZus{}num}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]} \PY{o}{=} \PY{n}{temp}\PY{p}{[}\PY{n}{l}\PY{p}{]}                
                    \PY{k}{else}\PY{p}{:}
                        \PY{k}{for} \PY{n}{l} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{l1}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{p}{:}
                            \PY{n}{loci\PYZus{}abs\PYZus{}num} \PY{o}{=} \PY{n}{l1}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]} \PY{o}{+} \PY{n}{l}
                            \PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{loci\PYZus{}abs\PYZus{}num}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]} \PY{o}{=} \PY{n}{temp}\PY{p}{[}\PY{n}{l}\PY{p}{]}                
                        \PY{k}{for} \PY{n}{l} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{l1}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n+nb}{len}\PY{p}{(}\PY{n}{temp}\PY{p}{)}\PY{p}{)}\PY{p}{:}
                            \PY{n}{loci\PYZus{}abs\PYZus{}num} \PY{o}{=} \PY{n}{l2}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]} \PY{o}{+} \PY{n}{l} \PY{o}{\PYZhy{}} \PY{n}{l1}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
                            \PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{loci\PYZus{}abs\PYZus{}num}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]} \PY{o}{=} \PY{n}{temp}\PY{p}{[}\PY{n}{l}\PY{p}{]}

\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
MIT
ALL
RL
CALL4
6053
6053
6053
6053

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}7}]:} \PY{c+c1}{\PYZsh{} Test of the above procedure, the absolute difference should be 0}
        \PY{n}{sum\PYZus{}diffs} \PY{o}{=} \PY{l+m+mi}{0}
        \PY{k}{for} \PY{n}{cell\PYZus{}index} \PY{o+ow}{in} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{randint}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n+nb}{len}\PY{p}{(}\PY{n}{data\PYZus{}names}\PY{p}{)}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{)}\PY{p}{:}
            \PY{n}{cell} \PY{o}{=} \PY{n}{data\PYZus{}names}\PY{p}{[}\PY{n}{cell\PYZus{}index}\PY{p}{]}
            \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{choice}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{)}\PY{p}{:}
                \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{choice}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{)}\PY{p}{:}
                    \PY{n}{i} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{argmin}\PY{p}{(}\PY{p}{[}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{]}\PY{p}{)}
                    \PY{k}{if} \PY{n}{i} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{:}
                        \PY{n}{c1} \PY{o}{=} \PY{n}{chr2}
                        \PY{n}{c2} \PY{o}{=} \PY{n}{chr1}
                    \PY{k}{else}\PY{p}{:}
                        \PY{n}{c1} \PY{o}{=} \PY{n}{chr1}
                        \PY{n}{c2} \PY{o}{=} \PY{n}{chr2}
                    \PY{k}{for} \PY{n}{loci} \PY{o+ow}{in} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{choice}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n}{lengths}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{c1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{)}\PY{p}{:}
                        \PY{n}{sum\PYZus{}diffs} \PY{o}{+}\PY{o}{=} \PY{n+nb}{int}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{sum}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{loci} \PY{o}{+} \PY{n}{lengths}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{c1}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{]}\PY{p}{[}\PY{n}{c2}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{c1}\PY{p}{]}\PY{p}{[}\PY{n}{c2}\PY{p}{]}\PY{p}{[}\PY{n}{loci}\PY{p}{]}\PY{p}{)}\PY{p}{)}\PY{p}{)}
        \PY{k}{print}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Absolute Diference: }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{sum\PYZus{}diffs}\PY{p}{)}\PY{p}{)}
        \PY{k}{print} \PY{n+nb}{len}\PY{p}{(}\PY{n}{loci\PYZus{}data}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIT}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Absolute Diference: 0
6054

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}8}]:} \PY{c+c1}{\PYZsh{}\PYZsh{} Saving files}
        \PY{c+c1}{\PYZsh{}for cell in data\PYZus{}names:}
        \PY{c+c1}{\PYZsh{}    for loci in range(num\PYZus{}loci):}
        \PY{c+c1}{\PYZsh{}            file\PYZus{}name = \PYZdq{}data/\PYZpc{}s\PYZus{}loci\PYZpc{}d\PYZdq{}\PYZpc{}(cell, loci)}
        \PY{c+c1}{\PYZsh{}            print(file\PYZus{}name)}
        \PY{c+c1}{\PYZsh{}            np.save(file\PYZus{}name, loci\PYZus{}data[cell][loci])}
\end{Verbatim}

    \subsection{Calculating Loci-wise
distances:}\label{calculating-loci-wise-distances}

Now we can compare contact maps with each other based on the
\emph{signature distance} measure. For each contact map and each pair of
cells, we can perform a t-test on the signature distances in order to
see whether the differences are 0 or not. In the following code blocks,
first we calculate the signature distances between each pair of cells
and then some visualizaitons demonstrate how cells are different.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}9}]:} \PY{n}{chr2count} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{n}{count2chr} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
        \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
            \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]} \PY{o}{=} \PY{n}{count}
                \PY{n}{count2chr}\PY{p}{[}\PY{n}{count}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
                \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}

        \PY{n}{cells} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{all}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{call4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
        \PY{n}{graphlets} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
            \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
            \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
                \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                    \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]} \PY{o}{=} \PY{n}{readMat}\PYZbs{}
                    \PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{/data/graphlets/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{.graphlets}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PYZbs{}
                     \PY{p}{(}\PY{n}{root}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{,} \PY{n}{cell}\PY{p}{)}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{astype}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}

        \PY{n}{distances} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{n}{actual\PYZus{}distances} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
        \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
            \PY{n}{distances}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
            \PY{n}{actual\PYZus{}distances}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

        \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
            \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                    \PY{k}{continue}
                \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{4}\PY{p}{)}\PY{p}{)}
                \PY{n}{distances}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}
                \PY{n}{actual\PYZus{}distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n+nb+bp}{None}\PY{p}{]} \PY{o}{*} \PY{p}{(}\PY{n}{count}\PY{p}{)}
                \PY{n}{actual\PYZus{}distances}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{n}{actual\PYZus{}distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}
                \PY{c+c1}{\PYZsh{}cell (i, j) will store correlation between orbital j in }
                \PY{c+c1}{\PYZsh{}ith chromosome}

        \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
            \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                    \PY{k}{continue}
                \PY{k}{print}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}
                \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                    \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                        \PY{n}{count} \PY{o}{=} \PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}
                        \PY{n}{temp} \PY{o}{=}  \PY{n}{row\PYZus{}wise\PYZus{}graphlet\PYZus{}distance}\PY{p}{(}\PY{n}{graphlets}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]}\PYZbs{}
                                        \PY{p}{,} \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]}\PY{p}{)}
                        \PY{n}{actual\PYZus{}distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{count}\PY{p}{]} \PY{o}{=} \PY{n}{temp}
                        \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{temp}\PY{p}{)}
                        \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanstd}\PY{p}{(}\PY{n}{temp}\PY{p}{,} \PY{n}{ddof}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                        \PY{n}{t0}\PY{p}{,} \PY{n}{pvalue} \PY{o}{=} \PY{n}{t\PYZus{}test}\PY{p}{(}\PY{n}{temp}\PY{p}{,} \PY{n}{one\PYZus{}sided}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
                        \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]} \PY{o}{=} \PY{n}{t0}
                        \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{3}\PY{p}{]} \PY{o}{=} \PY{n}{pvalue}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
('mit', 'rl')
('all', 'mit')
('all', 'rl')
('all', 'call4')
('call4', 'mit')
('call4', 'rl')

    \end{Verbatim}

    \subsubsection{All pair-wise signature
distances}\label{all-pair-wise-signature-distances}

The following is a graph comparing contact maps in terms of average
loci-loci distances. For each pair of cells \(A\) and \(B\),
\(\bar{d}^{\hspace{.2cm}A,B}_{ij}\) is calculated. As can be seen in the
third graph, which also shows errors of half standard deviation, we can
see that almost none of the graphlet distances are not within half a
standard deviation of zero.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}10}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{200}\PY{p}{,} \PY{l+m+mi}{90}\PY{p}{)}
         \PY{n}{myIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{myIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{myIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{distances}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{distances}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{n}{cell2}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{y} \PY{o}{=} \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{myIndices}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}
                 \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{o}{+}\PY{l+m+mi}{1}
                 \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
                 \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{)}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell2} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                     \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}        
                     \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{cell2}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                         \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,}  \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}       
                 \PY{c+c1}{\PYZsh{}ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(cell1, cell2))}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mf}{1.5}\PY{p}{]}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Average loci\PYZhy{}loci distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Contact maps of chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s1}{\PYZsq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xticks}\PY{p}{(}\PY{n}{x}\PY{p}{)}

         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:20: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_15_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \section{MIC Analysis}\label{mic-analysis}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}11}]:} \PY{n}{header} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{MIT,ALL,RL,CALL4}\PY{l+s+s2}{\PYZdq{}}
         \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                 \PY{k}{for} \PY{n}{orbit} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}\PY{p}{:}
                     \PY{n}{filename} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{/data/orbits/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}orbit}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{.csv}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{root}\PY{p}{,} \PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{,} \PY{n}{orbit}\PY{p}{)}
                     \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{100000000}
                     \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
                         \PY{n}{cell} \PY{o}{=} \PY{n}{cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}
                         \PY{n}{n} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{[}\PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{o}{.}\PY{n}{lower}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{orbit}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{n}{n}\PY{p}{]}\PY{p}{)}
                     \PY{n}{temp} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{n}{n}\PY{p}{,} \PY{n+nb}{len}\PY{p}{(}\PY{n}{data\PYZus{}names}\PY{p}{)}\PY{p}{)}\PY{p}{)}
                     \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
                     \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{data\PYZus{}names}\PY{p}{:}
                         \PY{n}{temp}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{count}\PY{p}{]} \PY{o}{=} \PY{n}{graphlets}\PY{p}{[}\PY{n}{cell}\PY{o}{.}\PY{n}{lower}\PY{p}{(}\PY{p}{)}\PY{p}{]}\PY{p}{[}\PY{n}{chr1}\PY{p}{]}\PY{p}{[}\PY{n}{chr2}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{n}{n}\PY{p}{,} \PY{n}{orbit}\PY{p}{]}
                         \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                     \PY{n}{np}\PY{o}{.}\PY{n}{savetxt}\PY{p}{(}\PY{n}{filename}\PY{p}{,} \PY{n}{temp}\PY{p}{,} \PY{n}{delimiter}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{,}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{header}\PY{o}{=}\PY{n}{header}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}12}]:} \PY{c+c1}{\PYZsh{}for chr1 in range(1, 24):}
         \PY{c+c1}{\PYZsh{}    for chr2 in range(chr1, 24):}
         \PY{c+c1}{\PYZsh{}        for orbit in range(73):}
         \PY{c+c1}{\PYZsh{}            file\PYZus{}name = \PYZdq{}data/chr\PYZpc{}02d\PYZus{}chr\PYZpc{}02d\PYZus{}orbit\PYZpc{}02d.csv\PYZdq{}\PYZpc{}(chr1, chr2, orbit)}
         \PY{c+c1}{\PYZsh{}            command = \PYZdq{}java \PYZhy{}jar mine.jar \PYZpc{}s \PYZhy{}allPairs\PYZdq{}\PYZpc{}file\PYZus{}name}
         \PY{c+c1}{\PYZsh{}            print(command)}
         \PY{c+c1}{\PYZsh{}            os.system(command)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}13}]:} \PY{n}{cells} \PY{o}{=} \PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{all}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{call4}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}
         \PY{n}{chr2count} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{n}{count2chr} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                 \PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]} \PY{o}{=} \PY{n}{count}
                 \PY{n}{count2chr}\PY{p}{[}\PY{n}{count}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}
                 \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}

         \PY{n}{mics} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{n}{mics}\PY{p}{[}\PY{n}{cell}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{n}{count}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}\PY{p}{)}
                 \PY{n}{mics}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}
                 \PY{c+c1}{\PYZsh{}cell (i, j) will store correlation between orbital j in }
                 \PY{c+c1}{\PYZsh{}ith chromosome}

         \PY{c+c1}{\PYZsh{}for cell1 in cells:}
         \PY{c+c1}{\PYZsh{}    for cell2 in cells:}
         \PY{c+c1}{\PYZsh{}        if cell2 \PYZlt{}= cell1:}
         \PY{c+c1}{\PYZsh{}            continue}
         \PY{c+c1}{\PYZsh{}        print(cell1, cell2)}
         \PY{c+c1}{\PYZsh{}        count = 0 }
         \PY{c+c1}{\PYZsh{}        for chr1 in range(1, 24):}
         \PY{c+c1}{\PYZsh{}            for chr2 in range(chr1, 24):}
         \PY{c+c1}{\PYZsh{}                mics[cell1][cell2][count, :] =  row\PYZus{}wise\PYZus{}pearson(graphlets[cell1][chr1][chr2].T\PYZbs{}}
         \PY{c+c1}{\PYZsh{}                                , graphlets[cell2][chr1][chr2].T)}
         \PY{c+c1}{\PYZsh{}                count+=1}
         \PY{k+kn}{import} \PY{n+nn}{string}
         \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                 \PY{k}{for} \PY{n}{orbit} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{73}\PY{p}{)}\PY{p}{:}
                     \PY{n}{filename} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{../data/mics/chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}chr}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{\PYZus{}orbit}\PY{l+s+si}{\PYZpc{}02d}\PY{l+s+s2}{.csv,allpairs,cv=0.0,B=n\PYZca{}0.6,Results.csv}\PY{l+s+s2}{\PYZdq{}}\PYZbs{}
                     \PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{,} \PY{n}{orbit}\PY{p}{)}
                     \PY{k}{print}\PY{p}{(}\PY{n}{filename}\PY{p}{)}
                     \PY{n}{csvfile} \PY{o}{=} \PY{n+nb}{open}\PY{p}{(}\PY{n}{filename}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{r}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
                     \PY{n}{csvfile}\PY{o}{.}\PY{n}{readline}\PY{p}{(}\PY{p}{)}
                     \PY{k}{for} \PY{n}{line} \PY{o+ow}{in} \PY{n}{csvfile}\PY{p}{:}
                         \PY{n}{splitted} \PY{o}{=} \PY{n}{line}\PY{o}{.}\PY{n}{split}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{,}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
                         \PY{n}{a} \PY{o}{=} \PY{n}{string}\PY{o}{.}\PY{n}{replace}\PY{p}{(}\PY{n}{splitted}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZsh{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{lower}\PY{p}{(}\PY{p}{)}
                         \PY{n}{b} \PY{o}{=} \PY{n}{string}\PY{o}{.}\PY{n}{replace}\PY{p}{(}\PY{n}{splitted}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZsh{} }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}\PY{o}{.}\PY{n}{lower}\PY{p}{(}\PY{p}{)}
                         \PY{n}{c} \PY{o}{=} \PY{n+nb}{float}\PY{p}{(}\PY{n}{splitted}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
                         \PY{c+c1}{\PYZsh{}print(a, b, c)}
                         \PY{n}{mics}\PY{p}{[}\PY{n}{a}\PY{p}{]}\PY{p}{[}\PY{n}{b}\PY{p}{]}\PY{p}{[}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{,} \PY{n}{orbit}\PY{p}{]} \PY{o}{=} \PY{n}{c}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}14}]:} \PY{n}{numPairs} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{cellPair2num} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{:}
             \PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell2}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]} \PY{o}{=} \PY{n}{numPairs}
                 \PY{n}{numPairs} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{vectorMic} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{zeros}\PY{p}{(}\PY{p}{(}\PY{n}{numPairs}\PY{p}{,} \PY{n}{count}\PY{p}{,} \PY{l+m+mi}{73}\PY{p}{)}\PY{p}{)}
         \PY{k}{print}\PY{p}{(}\PY{n}{vectorMic}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
         \PY{n}{pairCount} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell2}\PY{p}{:}
                     \PY{k}{continue}        
                 \PY{n}{vectorMic}\PY{p}{[}\PY{n}{pairCount}\PY{p}{]} \PY{o}{=} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}
                 \PY{n}{pairCount} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{k}{print}\PY{p}{(}\PY{n}{numPairs}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 276, 73)
6

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}16}]:} \PY{n}{sameIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{n}{sameIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{c+c1}{\PYZsh{}print(\PYZdq{}Same indices: \PYZpc{}s\PYZbs{}n\PYZdq{}\PYZpc{}sameIndices)}
         \PY{n}{differentIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr1} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{chr1}\PY{o}{+}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
                 \PY{n}{differentIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{c+c1}{\PYZsh{}print(\PYZdq{}Different indices: \PYZpc{}s\PYZbs{}n\PYZdq{}\PYZpc{}differentIndices)}

         \PY{n}{indices} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[} \PY{l+m+mi}{0}\PY{p}{,}  \PY{l+m+mi}{1}\PY{p}{,}  \PY{l+m+mi}{2}\PY{p}{,}  \PY{l+m+mi}{4}\PY{p}{,}  \PY{l+m+mi}{5}\PY{p}{,}  \PY{l+m+mi}{6}\PY{p}{,}  \PY{l+m+mi}{7}\PY{p}{,}  \PY{l+m+mi}{8}\PY{p}{,} \PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{16}\PY{p}{,} \PY{l+m+mi}{17}\PY{p}{,} \PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{19}\PY{p}{,} \PY{l+m+mi}{20}\PY{p}{,} \PY{l+m+mi}{21}\PY{p}{,} \PY{l+m+mi}{22}\PY{p}{,} \PY{l+m+mi}{23}\PY{p}{,}
                 \PY{l+m+mi}{35}\PY{p}{,} \PY{l+m+mi}{36}\PY{p}{,} \PY{l+m+mi}{37}\PY{p}{,} \PY{l+m+mi}{38}\PY{p}{,} \PY{l+m+mi}{49}\PY{p}{,} \PY{l+m+mi}{50}\PY{p}{]}\PY{p}{)} 
         \PY{n}{nonIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{73}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{i} \PY{o+ow}{not} \PY{o+ow}{in} \PY{n}{indices}\PY{p}{:}
                 \PY{n}{nonIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{i}\PY{p}{)}

         \PY{c+c1}{\PYZsh{}print(\PYZdq{}Indices: \PYZpc{}s\PYZbs{}n\PYZdq{}\PYZpc{}indices)}
         \PY{c+c1}{\PYZsh{}print(\PYZdq{}Non\PYZhy{}indices: \PYZpc{}s\PYZdq{}\PYZpc{}nonIndices)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}17}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{200}\PY{p}{,} \PY{l+m+mi}{90}\PY{p}{)}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{k}{print}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
         \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cellPair2num}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{n}{tmp} \PY{o}{=} \PY{n}{vectorMic} \PY{o}{\PYZhy{}} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{vectorMic}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n}{keepdims}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
                 \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{tmp}\PY{p}{[}\PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{p}{:}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{indices}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}

                 \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
                 \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
                 \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                     \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                     \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                         \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Contact Maps of Chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Average MIC}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{c+c1}{\PYZsh{}ax.set\PYZus{}xticks(x)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(276, 73)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:21: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_22_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}18}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{200}\PY{p}{,} \PY{l+m+mi}{90}\PY{p}{)}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{k}{print}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
         \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cellPair2num}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{n}{tmp} \PY{o}{=} \PY{n}{vectorMic} \PY{o}{\PYZhy{}} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{vectorMic}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{p}{,} \PY{n}{keepdims}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
                 \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{tmp}\PY{p}{[}\PY{n}{cellPair2num}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{p}{:}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{p}{:}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}

                 \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
                 \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
                 \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                     \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                     \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                         \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Contact Maps of Chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Average MIC}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xticks}\PY{p}{(}\PY{n}{x}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(276, 73)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:21: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_23_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}19}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{200}\PY{p}{,} \PY{l+m+mi}{90}\PY{p}{)}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{k}{print}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
         \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{x} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{rl}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{p}{)}\PY{p}{)}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell} \PY{o}{\PYZlt{}} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{n}{custIndices}\PY{p}{,} \PY{p}{:}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{indices}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
                 \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)} \PY{o}{+} \PY{l+m+mi}{1}
                 \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
                 \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
                 \PY{k}{if} \PY{n}{cell1} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                     \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                     \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                         \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Contact Maps of Chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{Average MIC}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xticks}\PY{p}{(}\PY{n}{x}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(276, 73)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:21: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_24_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}20}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{legends} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{chosen\PYZus{}cell} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell} \PY{o+ow}{in} \PY{n}{mics}\PY{p}{[}\PY{n}{chosen\PYZus{}cell}\PY{p}{]}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell} \PY{o}{\PYZlt{}} \PY{n}{chosen\PYZus{}cell}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{mics}\PY{p}{[}\PY{n}{chosen\PYZus{}cell}\PY{p}{]}\PY{p}{[}\PY{n}{cell}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{n}{indices}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
                 \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)}
                 \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
                 \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
                 \PY{k}{if} \PY{n}{chosen\PYZus{}cell} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}} \PY{o+ow}{or} \PY{n}{cell} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                     \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                     \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1} 
                 \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{n}{tmp}\PY{p}{,} \PY{o}{=} \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chosen\PYZus{}cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{cell}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                                \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}
                 \PY{c+c1}{\PYZsh{}tmp, = ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(chosen\PYZus{}cell, cell))}
                 \PY{n}{legends}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{tmp}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Orbit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
                 \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{MIC}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{25}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{handles}\PY{o}{=}\PY{n}{legends}\PY{p}{,} \PY{n}{loc}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:12: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.
  if sys.path[0] == '':

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_25_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsection{MIC statsitical analysis}\label{mic-statsitical-analysis}

    \subsubsection{Comparison between contact
maps}\label{comparison-between-contact-maps}

    \subsubsection{Comparison between contact maps only considering non-zero
orbits}\label{comparison-between-contact-maps-only-considering-non-zero-orbits}

    \subsubsection{Comparison between
orbits}\label{comparison-between-orbits}

    \subsubsection{Comparison between orbits only for intra-chromosomal
contact
maps}\label{comparison-between-orbits-only-for-intra-chromosomal-contact-maps}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}21}]:} \PY{n}{numData} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{datashape} \PY{o}{=} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
                 \PY{n}{numData} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{axis} \PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{xIndices} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{276}\PY{p}{)}
         \PY{n}{yIndices} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{73}\PY{p}{)}
         \PY{n}{yIndices} \PY{o}{=} \PY{n}{indices}
         \PY{n}{averages} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{counts} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{data} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}        
                 \PY{n}{tag} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}
                 \PY{n}{tmp} \PY{o}{=} \PY{n}{mics}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{n}{xIndices}\PY{p}{,} \PY{p}{:}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{n}{yIndices}\PY{p}{]}
                 \PY{n}{averages}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{tmp}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{axis}\PY{p}{)}\PY{p}{)}
                 \PY{n}{counts}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{tmp}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{axis}\PY{p}{]}\PY{p}{)}           
                 \PY{k}{if} \PY{n}{axis} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{:}
                     \PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{tmp}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{tmp}\PY{o}{.}\PY{n}{T}
                 \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{averages} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{averages}\PY{p}{)}
         \PY{n}{counts} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{counts}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{o}{.}\PY{n}{reshape}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{k}{print}\PY{p}{(}\PY{n}{averages}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 23)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}22}]:} \PY{k}{print}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}
         \PY{n}{manova}\PY{p}{(}\PY{n}{data}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{wilk}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
['all vs rl', 'call4 vs mit', 'mit vs rl', 'all vs call4', 'all vs mit', 'call4 vs rl']
Method Used: WILK
('n: ', 1656.0)
('k: ', 23)
('m: ', 6)
('a: ', 1640.5)
('b\_num', 13221.0)
('b\_denom', 549.0)
('b: ', 4.907338098512752)
('c: ', 56.5)
('gamma: ', 0.5654068065417478)
('s', 5)
('t', 8.5)
('u', 813.0)
('df1', 115)
('df2', 7993.988150610169)
('F', 8.565054626003658)
('alpha:', 0.05)
('F-crit:', 1.2284905642879855)
('p-value:', 1.1711969397971584e-128)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}23}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{i} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{data}\PY{p}{:}
             \PY{n}{i} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{k}{print}\PY{p}{(}\PY{n}{averages}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
             \PY{n}{means} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{averages}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
             \PY{n}{y} \PY{o}{=} \PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{nanmean}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{means}\PY{p}{)}
             \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)}
             \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
             \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{)}
             \PY{k}{if} \PY{n}{tag}\PY{o}{.}\PY{n}{find}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{:}
                 \PY{n}{color} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                 \PY{n}{mit\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{color} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}        
                 \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{tag}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                     \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,}  \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{)}       
             \PY{c+c1}{\PYZsh{}ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(cell1, cell2))}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}1.5, 1.5])}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Average loci\PYZhy{}loci distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{12}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Contact maps of chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s1}{\PYZsq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{12}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Corrected Average MIC}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Contact maps of chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s1}{\PYZsq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Orbits}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}    
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}xticks(x)}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}.2, .2])}

         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 23)
(6, 23)
(6, 23)
(6, 23)
(6, 23)
(6, 23)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:20: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_33_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}24}]:} \PY{n}{newData} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{p}{[}\PY{p}{]}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{cancer vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{p}{[}\PY{p}{]}\PY{p}{\PYZcb{}}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{data}\PY{p}{:}
             \PY{k}{if} \PY{n}{tag}\PY{o}{.}\PY{n}{find}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{:}
                 \PY{n}{newData}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{newData}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{cancer vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{newData}\PY{p}{:}
             \PY{k}{print}\PY{p}{(}\PY{n}{tag}\PY{p}{)}
             \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
             \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{o}{.}\PY{n}{reshape}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
             \PY{k}{print}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
normal vs cancer
(828, 23)
cancer vs cancer
(828, 23)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}25}]:} \PY{k}{print}\PY{p}{(}\PY{n}{newData}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}
         \PY{n}{manova}\PY{p}{(}\PY{n}{newData}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{wilk}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
['normal vs cancer', 'cancer vs cancer']
Method Used: WILK
('n: ', 1656.0)
('k: ', 23)
('m: ', 2)
('a: ', 1642.5)
('b\_num', 525.0)
('b\_denom', 525.0)
('b: ', 1.0)
('c: ', 10.5)
('gamma: ', 0.8137462912026895)
('s', 1)
('t', 10.5)
('u', 815.0)
('df1', 23)
('df2', 1632.0)
('F', 16.240830195044396)
('alpha:', 0.05)
('F-crit:', 1.5358774221121692)
('p-value:', 7.068138545917349e-58)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}26}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{25}\PY{p}{,} \PY{l+m+mi}{9}\PY{p}{)}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{i} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{newData}\PY{p}{:}
             \PY{n}{i} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{k}{print}\PY{p}{(}\PY{n}{averages}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
             \PY{n}{means} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{averages}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
             \PY{n}{y} \PY{o}{=} \PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{means}\PY{p}{)}
             \PY{n}{y\PYZus{}err} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{o}{.}\PY{l+m+mi}{1}
             \PY{n}{x} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n+nb}{range}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{y}\PY{p}{)}\PY{p}{)}\PY{p}{)}
             \PY{n}{xnew} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{n}{x}\PY{o}{.}\PY{n}{min}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{n}{x}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,}\PY{l+m+mi}{500}\PY{p}{)} \PY{c+c1}{\PYZsh{}300 represents number of points to make between T.min and T.max}
             \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{spline}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,}\PY{n}{xnew}\PY{p}{)}
             \PY{k}{if} \PY{n}{tag} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
             \PY{n}{ax}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{ }\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{tag}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                     \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{4}\PY{p}{,}  \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color1}\PY{p}{)}
             \PY{c+c1}{\PYZsh{}ax.errorbar(x, y, yerr=y\PYZus{}err, linestyle=\PYZdq{}dashed\PYZdq{}, marker=\PYZdq{}None\PYZdq{}, color=color2)}
             \PY{c+c1}{\PYZsh{}ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(cell1, cell2))}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}1.5, 1.5])}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Corrected Average MIC}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Contact maps of chromosome }\PY{l+s+si}{\PYZpc{}d}\PY{l+s+s1}{\PYZsq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{chr1}\PY{p}{)}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Contact Maps}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}xticks(x)}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}.3, .3])}

         \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{prop}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{size}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{l+m+mi}{15}\PY{p}{\PYZcb{}}\PY{p}{)}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 23)
(6, 23)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
/home/bzr0014/git/watson/virt/lib/python2.7/site-packages/ipykernel\_launcher.py:21: DeprecationWarning: `spline` is deprecated!
spline is deprecated in scipy 0.19.0, use Bspline class instead.

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_36_2.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}31}]:} \PY{n}{numData} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}
                 \PY{n}{datashape} \PY{o}{=} \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{o}{.}\PY{n}{shape}
                 \PY{c+c1}{\PYZsh{}print(datashape)}
                 \PY{n}{numData} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{axis} \PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{xIndices} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{276}\PY{p}{)}
         \PY{n}{yIndices} \PY{o}{=} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{yIndices} \PY{o}{=} \PY{n}{indices}
         \PY{n}{averages} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{counts} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{n}{data} \PY{o}{=} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
         \PY{n}{count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{cell1} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
             \PY{k}{for} \PY{n}{cell2} \PY{o+ow}{in} \PY{n}{cells}\PY{p}{:}
                 \PY{k}{if} \PY{n}{cell2} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{cell1}\PY{p}{:}
                     \PY{k}{continue}        
                 \PY{n}{tag} \PY{o}{=} \PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{ vs }\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{cell1}\PY{p}{,} \PY{n}{cell2}\PY{p}{)}
                 \PY{n}{tmp} \PY{o}{=} \PY{n}{distances}\PY{p}{[}\PY{n}{cell1}\PY{p}{]}\PY{p}{[}\PY{n}{cell2}\PY{p}{]}\PY{p}{[}\PY{p}{:}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{o}{.}\PY{n}{reshape}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
                 \PY{n}{averages}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{tmp}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{axis}\PY{p}{)}\PY{p}{)}
                 \PY{n}{counts}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{tmp}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{1}\PY{o}{\PYZhy{}}\PY{n}{axis}\PY{p}{]}\PY{p}{)}           
                 \PY{k}{if} \PY{n}{axis} \PY{o}{==} \PY{l+m+mi}{1}\PY{p}{:}
                     \PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{tmp}
                 \PY{k}{else}\PY{p}{:}
                     \PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{tmp}\PY{o}{.}\PY{n}{T}
                 \PY{n}{count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
         \PY{n}{averages} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{averages}\PY{p}{)}
         \PY{n}{counts} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{counts}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{uint32}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}\PY{o}{.}\PY{n}{reshape}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{c+c1}{\PYZsh{}print(averages.shape)}
         \PY{n}{manova}\PY{p}{(}\PY{n}{data}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{wilk}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Method Used: WILK
('n: ', 1656.0)
('k: ', 1)
('m: ', 6)
('a: ', 1651.5)
('b\_num', 21.0)
('b\_denom', 21.0)
('b: ', 1.0)
('c: ', 1.5)
('gamma: ', 0.9111131405886976)
('s', 1)
('t', 1.5)
('u', 824.0)
('df1', 5)
('df2', 1650.0)
('F', 32.19431517229249)
('alpha:', 0.05)
('F-crit:', 2.219521132720305)
('p-value:', 2.1404345664824635e-31)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}32}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{3}\PY{p}{)}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{i} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{data}\PY{p}{:}
             \PY{n}{i} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{k}{print}\PY{p}{(}\PY{n}{averages}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
             \PY{n}{means} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{averages}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
             \PY{n}{x} \PY{o}{=} \PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{means}\PY{p}{)}
             \PY{n}{x\PYZus{}err} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{o}{.}\PY{l+m+mo}{05}
             \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones\PYZus{}like}\PY{p}{(}\PY{n}{x}\PY{p}{)} \PY{o}{*} \PY{n}{i}\PY{o}{*}\PY{o}{.}\PY{l+m+mi}{1}
             \PY{n}{xnew} \PY{o}{=} \PY{n}{x}\PY{c+c1}{\PYZsh{}np.linspace(x.min(),x.max(),500) \PYZsh{}300 represents number of points to make between T.min and T.max}
             \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{y}\PY{c+c1}{\PYZsh{}spline(x, y,xnew)}
             \PY{k}{if} \PY{n}{tag}\PY{o}{.}\PY{n}{find}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{n}{mit\PYZus{}count}\PY{p}{]}
                 \PY{n}{mit\PYZus{}count}\PY{o}{+}\PY{o}{=}\PY{l+m+mi}{1}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{n}{cancer\PYZus{}count}\PY{p}{]}
                 \PY{n}{cancer\PYZus{}count} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{tag}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                     \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{5}\PY{p}{,}  \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color1}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{xerr}\PY{o}{=}\PY{n}{x\PYZus{}err}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dashed}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{None}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color1}\PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{5}\PY{p}{)}
             \PY{c+c1}{\PYZsh{}ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(cell1, cell2))}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}1.5, 1.5])}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylabel(\PYZsq{}Corrected Average MIC\PYZsq{}, fontsize=20)}
         \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Corrected Average Distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
         \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}yticks}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{data}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{o}{.}\PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}yticklabels}\PY{p}{(}\PY{n}{data}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}
         \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{o}{.}\PY{l+m+mi}{12}\PY{p}{,} \PY{o}{.}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}

         \PY{c+c1}{\PYZsh{}plt.legend(prop=\PYZob{}\PYZsq{}size\PYZsq{}:15\PYZcb{})}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 1)
(6, 1)
(6, 1)
(6, 1)
(6, 1)
(6, 1)

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_38_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}33}]:} \PY{n}{newData} \PY{o}{=} \PY{p}{\PYZob{}}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{p}{[}\PY{p}{]}\PY{p}{,} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{cancer vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}\PY{p}{[}\PY{p}{]}\PY{p}{\PYZcb{}}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{data}\PY{p}{:}
             \PY{k}{if} \PY{n}{tag}\PY{o}{.}\PY{n}{find}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{mit}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mi}{0}\PY{p}{:}
                 \PY{n}{newData}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{newData}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{cancer vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{newData}\PY{p}{:}
             \PY{c+c1}{\PYZsh{}print(tag)}
             \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{)}
             \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]} \PY{o}{=} \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{o}{.}\PY{n}{reshape}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{o}{.}\PY{n}{shape}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{)}
             \PY{c+c1}{\PYZsh{}print(newData[tag].shape)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}34}]:} \PY{k}{print}\PY{p}{(}\PY{n}{newData}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}
         \PY{n}{manova}\PY{p}{(}\PY{n}{newData}\PY{p}{,} \PY{n}{method}\PY{o}{=}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{hotelling}\PY{l+s+s1}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
['normal vs cancer', 'cancer vs cancer']
Method Used: HOTELLING
('n: ', 1656.0)
('k: ', 1)
('m: ', 2)
('s', 1)
('t', -0.5)
('u', 826.0)
('T20: ', 0.012390436681119634)
('df1', 1.0)
('df2', 1654.0)
('F', 20.493782270571874)
('alpha:', 0.05)
('F-crit:', 3.8470871354481613)
('p-value:', 6.409969039925188e-06)

    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}35}]:} \PY{n}{pylab}\PY{o}{.}\PY{n}{rcParams}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{figure.figsize}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{(}\PY{l+m+mi}{15}\PY{p}{,} \PY{l+m+mi}{3}\PY{p}{)}
         \PY{n}{chr1} \PY{o}{=} \PY{l+m+mi}{14}
         \PY{n}{custIndices} \PY{o}{=} \PY{p}{[}\PY{p}{]}
         \PY{k}{for} \PY{n}{chr2} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{24}\PY{p}{)}\PY{p}{:}
             \PY{k}{if} \PY{n}{chr1} \PY{o}{\PYZlt{}} \PY{n}{chr2}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr1}\PY{p}{,} \PY{n}{chr2}\PY{p}{)}\PY{p}{]}\PY{p}{)}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{custIndices}\PY{o}{.}\PY{n}{append}\PY{p}{(}\PY{n}{chr2count}\PY{p}{[}\PY{p}{(}\PY{n}{chr2}\PY{p}{,} \PY{n}{chr1}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{n}{ax} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{subplot}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
         \PY{n}{i} \PY{o}{=} \PY{o}{\PYZhy{}}\PY{l+m+mi}{1}
         \PY{n}{mit\PYZus{}count} \PY{o}{=} \PY{n}{cancer\PYZus{}count} \PY{o}{=} \PY{l+m+mi}{0}
         \PY{k}{for} \PY{n}{tag} \PY{o+ow}{in} \PY{n}{newData}\PY{p}{:}
             \PY{n}{i} \PY{o}{+}\PY{o}{=} \PY{l+m+mi}{1}
             \PY{k}{print}\PY{p}{(}\PY{n}{averages}\PY{o}{.}\PY{n}{shape}\PY{p}{)}
             \PY{n}{means} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{averages}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}
             \PY{n}{x} \PY{o}{=} \PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{\PYZhy{}} \PY{n}{means}\PY{p}{)}
             \PY{n}{x\PYZus{}err} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{n}{newData}\PY{p}{[}\PY{n}{tag}\PY{p}{]}\PY{p}{,} \PY{n}{axis}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)} \PY{o}{*} \PY{o}{.}\PY{l+m+mi}{1}
             \PY{n}{y} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{ones\PYZus{}like}\PY{p}{(}\PY{n}{x}\PY{p}{)} \PY{o}{*} \PY{n}{i}\PY{o}{*}\PY{o}{.}\PY{l+m+mi}{1}
             \PY{n}{xnew} \PY{o}{=} \PY{n}{x}\PY{c+c1}{\PYZsh{}np.linspace(x.min(),x.max(),500) \PYZsh{}300 represents number of points to make between T.min and T.max}
             \PY{n}{y\PYZus{}smooth} \PY{o}{=} \PY{n}{y}\PY{c+c1}{\PYZsh{}spline(x, y,xnew)}
             \PY{k}{if} \PY{n}{tag} \PY{o}{==} \PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{normal vs cancer}\PY{l+s+s1}{\PYZsq{}}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{mit\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}
             \PY{k}{else}\PY{p}{:}
                 \PY{n}{color1} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
                 \PY{n}{color2} \PY{o}{=} \PY{n}{cancer\PYZus{}colors}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}    
             \PY{c+c1}{\PYZsh{}ax.scatter(x, y)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{xnew}\PY{p}{,} \PY{n}{y\PYZus{}smooth}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{o}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+si}{\PYZpc{}s}\PY{l+s+s2}{\PYZdq{}}\PY{o}{\PYZpc{}}\PY{p}{(}\PY{n}{tag}\PY{o}{.}\PY{n}{upper}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PYZbs{}
                     \PY{p}{,} \PY{n}{linewidth}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{,}  \PY{n}{alpha}\PY{o}{=}\PY{o}{.}\PY{l+m+mi}{5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color1}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{errorbar}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{xerr}\PY{o}{=}\PY{n}{x\PYZus{}err}\PY{p}{,} \PY{n}{linestyle}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{dashed}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{None}\PY{l+s+s2}{\PYZdq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color1}\PY{p}{)}
             \PY{c+c1}{\PYZsh{}ax.plot(x, y, marker=\PYZdq{}.\PYZdq{}, label=\PYZdq{}\PYZpc{}s vs \PYZpc{}s\PYZdq{}\PYZpc{}(cell1, cell2))}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylim([\PYZhy{}1.5, 1.5])}
             \PY{c+c1}{\PYZsh{}ax.set\PYZus{}ylabel(\PYZsq{}Corrected Average MIC\PYZsq{}, fontsize=20)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{Corrected Average Distance}\PY{l+s+s1}{\PYZsq{}}\PY{p}{,} \PY{n}{fontsize}\PY{o}{=}\PY{l+m+mi}{20}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}yticks}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{n+nb}{len}\PY{p}{(}\PY{n}{newData}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{o}{.}\PY{l+m+mi}{1}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}yticklabels}\PY{p}{(}\PY{n}{newData}\PY{o}{.}\PY{n}{keys}\PY{p}{(}\PY{p}{)}\PY{p}{)}
             \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{o}{.}\PY{l+m+mo}{075}\PY{p}{,} \PY{o}{.}\PY{l+m+mo}{050}\PY{p}{]}\PY{p}{)}

         \PY{c+c1}{\PYZsh{}plt.legend(prop=\PYZob{}\PYZsq{}size\PYZsq{}:15\PYZcb{})}
         \PY{n}{plt}\PY{o}{.}\PY{n}{show}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
(6, 1)
(6, 1)

    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{output_41_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    % Add a bibliography block to the postdoc

    \end{document}